### 2013-03-01 0_frontmatter/abstract.tex
201303_Mar01

% Thesis Abstract -----------------------------------------------------

%\begin{abstractslong}    %uncommenting this line, gives a different abstract heading
\begin{abstracts}        %this creates the heading for the abstract page

Put your abstract or summary in 0\_frontmatter/abstract.tex.

\end{abstracts}
%\end{abstractlongs}

% ----------------------------------------------------------------------