### 2013-03-01 3/sar.tex

\chapter{Consistent Measures of Distance for Univariate SAR Data} %chapter 3
\label{chap:sar}

%\section{SAR speckle statistic model}
%\label{sec:heteroskdastic_model}

Useful and robust statistical estimators require un-assuming models. 
In this chapter a homoskedastic model for the usually heteroskedastic SAR is proposed. %speckle filtering.
In the first section the heteroskedastic nature of original amplitude and intensity SAR data will be explained. 
Relying solely on statistical principles,
  log-transformed SAR data is shown to be homoskedastic.
Together with the consistent sense of variance, a few consistent measures of distance are proposed for the traditional SAR data.

\section{Original Heteroskedastic Model of SAR data}

SAR speckle phenomena are explained as the interference of many coherent but dephased back-scaterring components;
  each reflecting from different and distributed elementary scaterers \citep{Oliver_ProcIEEE_1963, Leith_ProcIEEE_1971}. 
The random nature of the SAR data arises
  due to the unknown and varying location, height and thus distance of each elementary scaterer;
  which, in turn, generates random phases in their back-scattering responses.
Assuming the number of these elementary back scatterers is sufficiently large,
  then the Central Limit Theorem is applicable \citep{Goodman_Springer_1975}.
Consequently the interference of these randomly phased responses can be considered as
  a random walk on the 2D complex plane \citep{Goodman_JOptSocAm_76}.  
Specifically the real part $A_r$ as well as the imaginary part $A_i$ of the observed SAR signal $A$ can be considered as
  random variables from uncorrelated Gaussian distributed stochastic processes with zero means and indentical variances $\sigma^2/2$  \citep{Lee_CRCPress_2009}. 
Their probability density functions (PDF) are given as:

\begin{equation}
\label{eqn:SAR Real and Imaginary Components PDF}
\caption{eqn:SAR Real and Imaginary Components PDF}
pdf(A_x)=\frac{1}{\sqrt{\pi} \sigma} e^{\left( \frac{A_x^2}{\sigma^2} \right) }
\end{equation}

It can then be proved that the measurable amplitude $A=\sqrt{A_r^2+A_i^2}$ is a random variable of Rayleigh distribution
  and subsequently the SAR intensity $I=A^2=(A_r^2+A_i^2)$ is a random variable of a negative exponentially distributed random process.
That is:

\begin{eqnarray}
pdf(A) &=& \frac{2A}{\sigma^2}e^{ \left( -\frac{A^2}{\sigma^2} \right) }\\
pdf(I) &=& \frac{1}{\sigma^2}e^{\left( -\frac{I}{\sigma^2} \right) }
\end{eqnarray}

From a statistical viewpoint, the multiplicative nature of amplitude and intensity data can be explained as follows. 
Let us consider two fixed, independent to $\sigma$ unit distributions given below:

\begin{eqnarray}
pdf(A_1) &=& 2A_1 e^{ \left( A_1^2 \right) }\\
pdf(I_1) &=& e^{ \left( -I_1 \right) }
\end{eqnarray}

Using the variable change theorem,
%  it is then trivial to show that:
    it can be shown that:
    the SAR amplitude and intensity are equal to %are simply
    a scaled version of these unit variables,
    specifically $A= \sigma A_1 $ and $I= \sigma^2 I_1 $. 
These relationships evidently manifest a mutiplicative nature. 
In fact, the condition has long been noted,
  but from different perspectives,
  in various proposed SAR models (e.g. \cite{Jakeman_1980_JPhysAMathGen}).

Let us now consder the population expected mean and variance of the four distributions.
They are given in Table \ref{tbl:orginal_sar_avg_var}. 
If spatial homogeneity is defined as an imaging areas having the same back-scattering coefficient $\sigma$,
  then over a homogeneous scene, the measured values can then be considered as samples coming from a single stochastic process.
However, over heterogeneous areas where $\sigma$ varies significantly,
  it is then evidently clear that: the amplitude ($A$) as well as the intensity ($I$) of SAR data suffer from hetoroskedastic phenomena,
  which is defined as the dependence of conditional expected variance of SAR data on the conditional expectation of mean. 
In the context of speckle filtering which aims to estimate the unknown parameter $\sigma$,
  Table \ref{tbl:orginal_sar_avg_var} also clearly indicates the ambiguous circle.
That is, in the original domain, estimating variance is equal to estimating the mean
  and is as hard as the main problem itself.
  %, i.e. estimating the unknown parameter $\sigma$

\begin{table}[!h]
\normalsize
\centering
\begin{tabular}{|l|l|}
\hline
Mean & Variance \\
\hline
$avg(A_1) = \frac{ \sqrt{\pi}}{2}$ & $var(A_1) = \frac{(4-\pi)}{4}$ \\
$avg(I_1) = 1$ & $var(I_1) = 1$ \\
$avg(A) = \frac{\sqrt{\pi}}{2} \cdot \sigma $ & $var(A) = \frac{(4-\pi)}{4} \cdot \sigma^2 $ \\
$avg(I) = \sigma^2 $ & $ var(I) = \sigma^4$ \\
\hline
\end{tabular}
\caption{ Mean and Variance of Original SAR data }
\label{tbl:orginal_sar_avg_var}
\end{table}

The formulas above have long been noted in this research field. 
In fact while pioneering the estimation of the equivalent number of look (ENL) index, 
 it has been noted that the ratio of expected standard deviation to mean is a constant in both cases (i.e. $snr(A)=\sqrt{\frac{4}{\pi}-1}$ and $snr(I)=1$) \cite{Lee_1992_IMA_298}.
Here we just offer a different perspective,
  one that avoids the negative effects of heteroskedasticity in SAR data processing.
%Here, we just offer a different interpretation, which sets the stage for the discussion of logarithmic transformation.

\section{The Effects of Heteroskedasticity in SAR Data Processing}

%\subsection{Statistical Heteroskedasticity of SAR data}

The nature of SAR speckle is stochastic over homogeneous areas and heteroskedastic heterogeneously. 
Statistical models are used to describe how the underlying back-scattering coefficient ($\sigma$) affects the distribution of the observable SAR data. 
%For about 30 years, speckle filtering has been an active research area, with new methods being introduced steadily \cite{Lee_TGRS_2009}. 
%That is because
And while the statistical model within individual resolution cells is well established,
  the applicability of these models, however, are restricted to homogeneous areas. 
Practical images, nevertheless, are heterogeneous. 
Crucially it is this spatial variation that is of high interest. 
This fact gives rise to the question that seemed obvious:
  %how to call an analysis area heterogeneous and subsequently
  what to do in the case of heterogeneity.

Various statistical models for heterogeneous areas have been proposed (see \cite{Touzi_2002_TGRS} for a detailed review). 
Unfortunately, while most of the models highlight the multiplicative nature of sub-pixel or homogeneous original SAR data, in extending the model to heterogeneous images, virtually none has noted that spatial variation also gives rise to heteroskedasticity phenomena. 
Heteroskedasticity, as explained later, is defined as the dependence of conditional expected variance of original SAR data on the conditional expectation of its mean or equivalently its underlying back-scattering coefficient $(\sigma)$.

Speckle filtering, by and large, are estimators attempting to estimate the unknown coefficient $\sigma$ from the observable SAR signal $(A)$. 
Thus, SAR speckle-filtering can be and has been positioned within the context of estimation theory \cite{Touzi_2002_TGRS}. 
The statistical framework attempts to estimate unknown statistical parameters from the observable data. 
The stages consist of statistical modelling, estimator development and evaluating the performance of the statistical estimators. 
%In this section, the negative effects of heteroskedasticity on speckle filtering is analysed.
It is believed that heteroskedasticity, in its turn, give rise to serious negative impacts on various stages of speckle-filtering.

In modelling, heteroskedasticity has direct consequences in the central question of homogeneity or heterogeneity. 
In optical images, the contrast or variance among neighboring pixels have been used to measure homogeneity. 
Unfortunately such techniques do not appear to be effective under the heteroskedastic condition of original SAR data. 
In SAR images, both of these measures are dependent on the underlying signal $(\sigma)$. 
This make the problem of estimating variance equal to the problem of estimating the mean and is the same as the main problem: estimating $\sigma$. 
The fact give rise to the ambiguity circle in SAR speckle filtering.

A case in point can be found in the recently published improved sigma filter\cite{Lee_TGRS_2009}. 
The technique determines outlying points as being too far away from the standard deviation. 
However, as is done in \cite{Lee_TGRS_2009}, to estimate standard deviation, an estimator of the mean is required and used. 
It is interesting to note that the MMSE estimator used to estimate the mean in \cite{Lee_TGRS_2009}, itself alone, is a rather successful speckle-filter \cite{Lee_PAMI_1980}.

Heteroskedasticity also poses numerous challenges in designing an efficient estimator. Heteroskedasticity directly violates Gauss-Markov theorem's homoskedastic assumption. 
Thus it renders the efficiency of any naive Ordinary Least Square estimator in serious doubt. 
If the variance is known \textit{a priori}, it has been proven \cite{Aitken_1934_ProcsRoyalSocEdinburg}, that a weighted mean estimator is the best linear unbiased estimator. 
Interestingly, as noted by Lopes \cite{Lopes_TGRS_1990}, most known common successful adaptive filters \cite{Lee_PAMI_1980, Kuan_1985_PAMI, Frost_PAMI_1982} do make use of weighted mean estimations. 
The caveat however is that in SAR speckle filtering, variance is not known \textit{a priori}. 
And even though variance can be estimated from observable values, as the ambiguity circle goes, estimating the variance is as hard as estimating the underlying coefficient $\sigma$ itself.

Heteroskedasticity also affects the performance evaluation stage of such a framework.
Statistical estimators are routinely evaluated using their bias and variance performance,
  which when combined normally results in MSE evaluation.
Unfortunately, under heteroskedastic conditions, Least Squared Error estimation may not be optimal.   
In other words, Minimal Squared Error can no longer serves as a reliable evaluation metric.
A number of different evaluation metrics have been proposed for SAR speckle filters,
  however, none have appeared capable of taking on the MSE's previous role.

%Heteroskedasticity also decrease the effectiveness of estimators by increasing the variance of even the best possible estimations. 
%This fact makes it tricky to evaluate the performance of different estimators, even if quantitative experimental results are reported.
%In fact several quantitative simulated experiments has been carried out to compare the performance of different estimators\cite{Lee_TGRS_2009}. 
%However only results of single experiments were reported. 
%Due to the stochastic nature of the simulated experiment process, more experiments should be done to report the effectiveness, i.e. variance, of each estimators' results.

Last but certainly not least, is the bad impact of heteroskedasticity on SAR image interpretation. 
Most of the task to be carried out in interpreting SAR images almost invariably involve target detection, target segmentation and/or target classification. 
All of these tasks require good discriminant functions. 
Foundational to all of these is the need for a consistence sense of distance. 
Unfortunately, by definition, heteroskedasticity provides inconsistent measures of distance. 
This inconsistent sense of distance coupled with the failure of the ordinary least square regression methods are believed to have caused a large class of artificial neural networks as well as a number of other computational intelligence methods to under-perform on SAR data.

%In our work, we aim to neutralize these various negative impacts. 
%The paper is organized as follows. 
%Section \ref{sec:heteroskdastic_model} will be devoted to discussion of the heteroskedastic phenomena in the original per-pixel SAR amplitude and intensity model. 
%Logarithmic transformation is extended to the model and the homoskedastic property of it is asserted. 
%Section \ref{sec:variance_in_log_domain} will discuss how sampling distribution of variance in the log-transformed domain can be used to model various degrees of homogeneity. 
%Section \ref{sec:k_mle} will presents a novel speckle filtering algorithm developed from heterogenous inference made possible by statistical testing of sample variances.
%Section \ref{sec:k_mle_result} will present evaluation of our estimator both quantitatively through simulated experiments and qualitatively on real-life images.

\section{The Homoskedastic Effect of Logarithmic Transformation}

%\subsection{Homoskedastic effect of Logarithmic Transformation }

In this chapter, a base-2 logarithmic transformtion is proposed for SAR data. 
The choice of base-2 is due to implementation reasons
  since base-2 logarithm can be computed faster on digital computer than either natural or decimal logarithms,
  and yet the ability to transform heterskedactic speckle into a homoskedastic relationship is still maintained.
Under base-2 log-transformation, the original variables become:

\begin{eqnarray}
L_{1}^{A} &=& \log_2(A_1) = L_{1}^{I} / 2 \\
L_A &=& \log_2(A) 	= L_{1}^{A} + \log_2\sigma \\
L_{1}^{I} &=& \log_2(I_1) = 2 L_{1}^{A} \\
L_I &=& \log_2(I) 	= L_{1}^{I} + 2 \log_2\sigma
\end{eqnarray}

The equations above highlight the relationships among random variables in the log-transformed domain. 
Two conclusions become evident from these formulae. 
Firstly, in the log-transformed domain, working on either amplitude or intensity will tend to give identical results. 
Secondly, the effects of converting the multiplicative nature to an additive nature through logarithmic transformation
is clearly manifested. 
Now, bearing the relationship among the random variables in mind, it is then trivial to 
show that the probability distribution of these log-transformed variables are related as follows:

\begin{eqnarray}
pdf(L_{1}^{A}) &=& 2 \cdot 2^{\left( 2 L_{1}^{A} \right)}e^{-2^{\left( 2 L_{1}^{A} \right) }} \\
pdf(L_A) &=& 2 \cdot 2^{\left( 2 L_A - 2 \log_2 \sigma \right)}e^{-2^{\left( 2 L_A - 2 \log_2 \sigma \right)}} \\ 
pdf(L_{1}^{I}) &=& 2^{L_1^I} e^{-2^{L_1^I}}  \\
pdf(L_I) &=& 2^{\left( L_I - 2 \log_2 \sigma \right)} e^{-2^{\left( L_I - 2 \log_2 \sigma \right)} }  
%pdf(L_{1}^{A}) &=& 2 \cdot 2^{\left[ 2 L_{1}^{A} - 2^{2 L_{1}^{A}} \right]} \\
%pdf(L_A) &=& 2 \cdot 2^{\left[ \left( 2 L_A - 2 \log_2 \sigma \right) - 2^{\left( 2 L_A - 2 \log_2 \sigma \right)} \right]} \\ 
%pdf(L_{1}^{I}) &=& 2^{\left[ L_{1}^{I} - 2^{L_{1}^{I}} \right]} \\
%pdf(L_I) &=& 2^{\left[ \left( L_I - 2 \log_2 \sigma \right) - 2^{\left( L_I - 2 \log_2 \sigma \right)} \right]} \label{eqn:log_intensity_base_2_dist}
\end{eqnarray}

Noting that these distributions belong to the Fisher-Tippet family, the population expected mean and variances 
are obtained and presented in Table \ref{tbl:sar_log_domain_avg_var}, with $\gamma \approx 0.577$ being the Euler-Mascheroni constant. 
Most importantly, it can be seen that the means are biased and the variances are no longer related to $\sigma$

\begin{table}[h]
\normalsize
\centering
%\begin{center}

\begin{tabular}{|l|l|}
\hline
Mean & Variance \\
\hline
$avg(L_{1^A}) = \frac{ - \gamma }{2} \cdot \frac{1}{\ln2}$ & $var(L_{1^A}) = \frac{ \pi ^2}{24} \cdot \frac{1}{(\ln2)^2}$ \\
$avg(L_{1^I}) = - \gamma \cdot \frac{1}{\ln2} $ & $var(L_{1^I}) = \frac{ \pi ^2}{6} \cdot \frac{1}{(\ln2)^2} $ \\
$avg(L_A) = \frac{ - \gamma }{2} \cdot \frac{1}{\ln2} + \log_2{\sigma}$ & $var(L_A) = \frac{ \pi ^2}{24} \cdot \frac{1}{(\ln2)^2}$ \\
$avg(L_I) = - \gamma \cdot \frac{1}{\ln2} + 2 \log_2{\sigma}  $ & $ var(L_I) = \frac{ \pi ^2}{6} \cdot \frac{1}{(\ln2)^2}$ \\
\hline
\end{tabular}

\caption{ Mean and Variance of Log Transformed values: here the variance is no longer dependent on $\sigma$ }
\label{tbl:sar_log_domain_avg_var}
%\end{center}
\end{table}

Table \ref{tbl:sar_log_domain_avg_var} confirms the condition of homoskedasticity,  
defined as the independence of the conditional expected variance on the conditional expectation of the mean. 
This result is consistent with findings in \cite{Arsenault_JOptSocAm_1976}. 
The main difference would be the use of base-2 logarithm which is preferred here for more efficient computation.

Table \ref{tbl:sar_variables_properties} summarizes the discussion so far. 
It can be seen that while the original data, especially intensity values, should be preferred for 
multi-look processing, the log-transformed domain with its homoskedastic distribution offers consistent sense of variance.

\begin{table*}[t]
%\normalsize
%\footnotesize -2
\scriptsize
\centering

\begin{tabular}{|l|l|l|l|}
\hline
 RV & Relationships  & Variance (skedasticity) & Mean (biasness) \\
\hline
$A$ 
	& $A=\sigma A_1 $ 
	& Heteroskedastic $var(A) = \frac{(4-\pi)}{4} \cdot \sigma^2 $ 
	& Unbiased $avg(A) = \frac{\sqrt{\pi}}{2} \cdot \sigma $ \\
$I$ 
	& $I=A^2=\sigma^2 I_1 $ 
	& Heteroskedastic $ var(I) = \sigma^4$ 
	& Unbiased $avg(I) = \sigma^2 $\\
$L_A$ 
	& $L_A=\ln(A)=L_{1^A} + \log_2{\sigma}$ 
	& Homoskedastic $var(L_A) = \frac{ \pi ^2}{24} \cdot \frac{1}{(\ln2)^2}$ 
	& Biased $avg(L_A) = \frac{ - \gamma }{2} \cdot \frac{1}{\ln2} + \log_2{\sigma}$ \\
$L_I$ 
	& $L_I=\ln(I)=L_{1^I} + 2 \log_2{\sigma}$  
	& Homoskedastic $var(L_I) = \frac{ \pi ^2}{6} \cdot \frac{1}{(\ln2)^2}$ 
	& Biased $avg(L_I) = - \gamma \cdot \frac{1}{\ln2} + 2 \log_2{\sigma}  $ \\
\hline
\end{tabular}

\caption{ The properties of observable SAR random variables }
\label{tbl:sar_variables_properties}

\end{table*}

To experimentally validate this theoretical discussion, Fig. \ref{fig:modelled_response} plots the histogram of observable data within a 
known homogenous area (from a RADARSAT2 image) against modelled PDF response. 
An excellent agreement is observable in the graph which validate the log-transformed model. 
%In fact, this modelling has been used successful in explaining speckle phenomena, also verified 
%through scientific experiments \cite{Ulaby_TGRS_1988}.

\begin{figure}[!h]
\centering
%\centerline{
\begin{tabular}{c}
	\subfloat[amplitude]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/amplitude_histogram.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[intensity]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/intensity_histogram.eps} 	
		 \label{intensity}
	} \\
	\subfloat[log amplitude]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_amplitude_histogram.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[log intensity]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_intensity_histogram.eps} 	
		 \label{intensity}
	} 
\end{tabular}
%}
\caption{Observed histogram in homogenous area and modelled pdf response}
\label{fig:modelled_response}
\end{figure}

\section{Consistent Measures of Distance in the Log-Transformed Domain}

%\subsection{Consistent measures of distance in Log-Transformed domain}

The consistent feature is explored and illustrated in this section from two different perspectives. 
First, the back-scattering coefficient $\sigma$ is assumed to be \textit{known a priori}.
With that, let us consider the dispersion random 
variable defined as the distance between an observable sample and its expected value:

\begin{eqnarray}
D_A &=& A - avg(A) \\
D_I &=& I - avg(I) \\
D_{L^A} &=& L_A - avg(L_A) \\
D_{L^I} &=& L_I - avg(L_I) 
\end{eqnarray}

Noting the results from the previous section, the PDF for these variable can be derived as follows:

\begin{eqnarray}
pdf(D_A) &=& 2 \cdot \frac{\left( D_A + \sigma \sqrt{\pi}/2 \right)}{\sigma^2}e^{ \left[ - \frac{\left( D_A + \sigma \sqrt{\pi}/2 \right)^2}{\sigma^2}   \right] } \\
pdf(D_I) &=& \frac{1}{\sigma^2}e^{\left[ -\left( D_I + \sigma^2 \right) / \sigma^2 \right] } \\
pdf(D_{L^A}) &=& 2 \cdot 2^{\left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right)} e^{-2^{\left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right)}}\\
pdf(D_{L^I}) &=& 2^{\left( D_{L^I} + \frac{\gamma}{\ln2} \right)} e^{-2^{\left( D_{L^I} + \frac{\gamma}{\ln2} \right)}}
%pdf(D_{L^A}) &=& 2 \cdot 2^{\left[ \left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right) - 2^{\left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right)} \right]} \\
%pdf(D_{L^I}) &=& 2^{\left[ \left( D_{L^I} + \frac{\gamma}{\ln2} \right) - 2^{\left( D_{L^I} + \frac{\gamma}{\ln2} \right)} \right]}
\end{eqnarray}

From these equations, it is clear that these distributions, which are dependent on $\sigma$ in the original domain, 
are now independent of it when processed in the log-domain.

Let us investigate SAR data from the second perspective,
  where two adjacent resolution cells are known to have \textit{an identical but unknown} back-scattering coefficient $\sigma$.
Now, the contrast random variable is defined, also in the log-transformed domain, as the distance between two measured samples.

%From a second perspective, given two adjacent resolution cells that are known to have \textit{identical but unknown} back-scattering coefficient $\sigma$,
%  consider the contrast random variable defined, also in the log-transformed domain, as the distance between two 
%measured samples:

\begin{eqnarray}
C_A &=& A_1^\sigma - A_2^\sigma \\
C_I &=& I_1^\sigma - I_2^\sigma \\
C_{L^A} &=& L_1^{A_\sigma} - L_2^{A_\sigma} = log_2 { \left( {A_1}/{A_2} \right) }\\
C_{L^I} &=& L_1^{I_\sigma} - L_2^{I_\sigma} = log_2 { \left( {I_1}/{I_2} \right) }
\end{eqnarray}

Noting that $C_x = D_1^x - D_2^x$, it should come as no suprise that the measure of contrast is also consistent in the 
log-domain but inconsistent in the original domain. The PDF of the log-transformed variables can be expressed as:

\begin{eqnarray}
pdf(C_{L^A}) &=& 2 \frac{2^{\left(2 C_{L^A} \right)}}{\left[ 1+2^{\left( 2 C_{L^A} \right)} \right]^2} \ln2  \\
pdf(C_{L^I}) &=& \frac{2^{\left( C_{L^I} \right)}}{\left[ 1+2^{\left( C_{L^I} \right)} \right]^2} \ln2 
%pdf(C_{L^A}) &=& 2 \frac{2^{\left(2 C_{L^A} \right)}}{1+2^{\left( 2 C_{L^A} \right)}} \ln2  \\
%pdf(C_{L^I}) &=& \frac{2^{\left( C_{L^I} \right)}}{1+2^{\left( C_{L^I} \right)}} \ln2 
\end{eqnarray}

Evidently these distributions are also consistent, i.e. they are independent of $\sigma$.
To illustrate this property, Fig. \ref{fig:residual_as_noise} is generated using the data extracted from a homogeneous area in a RADARSAT-2 image.
As can be seen, it shows excellent agreement between the analytical PDF and the observable histogram of dispersion and contrast computed from an homogeneous area in the image. 
%To illustrate this property, Fig. \ref{fig:residual_as_noise} is generated by using data from the RADARSAT-2 
%image. As can be seen, it shows excellent agreement between the analytical pdf and the observable histogram of dispersion
%and contrast of an homogeneous area in the image. 
%Distance is calculated as the difference between each value point and the average value of that region, 
%while contrast is calculated as the difference between two horizontally adjacent values in the log-transformed domain.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[dispersion]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_intensity_dispersion_histogram.eps} 	
		 \label{amplitude}
	} 
	\hfill
	\subfloat[contrast]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_intensity_contrast_histogram.eps} 	
		 \label{intensity}
	}
\end{tabular}
\caption{Observed and modelled pdf of dispersion 
and contrast in homogenous log-transformed intensity images.}
\label{fig:residual_as_noise}
\end{figure}

Given that real SAR images are heterogenous and the back-scattering coefficient is unknown, it is evident that 
the measure of observable contrast in original SAR data differs, i.e. is inconsistent, across different homogenous 
areas. On the other hand, the observable contrast in the log-transformed domain is consistently the same across 
different homogenous areas. As such, one possible benefit is that should the sigma filter \cite{Lee_TGRS_2009} be 
designed against the $pdf(C_{L^I})$, then the scale estimator would probably no longer be required.

\subsection{Sampling distribution of Variance in Log-Transformed domain}

From the previous section result, two sample variance distribution ($V_x = C_x^2$) can be given analytically as

\begin{eqnarray}
pdf(V_{L^I}) &=& 
	\frac{\ln2}{2} \frac{1}{\sqrt{V_{L^I}}} \left( \frac{2^{\sqrt{V_{L^I}}}}{\left( 1+2^{\sqrt{V_{L^I}}} \right)^2} + \frac{2^{-\sqrt{V_{L^I}}}}{\left( 1+2^{-\sqrt{V_{L^I}}} \right)^2} \right)\\
pdf(V_{L^A}) &=&
	\frac{\ln2}{4} \frac{1}{\sqrt{V_{L^A}}} \left( \frac{2^{\sqrt{4 V_{L^A}}}}{\left( 1+2^{\sqrt{4 V_{L^A}}} \right)^2} + \frac{2^{-\sqrt{4 V_{L^A}}}}{\left( 1+2^{-2\sqrt{V_{L^A}}} \right)^2} \right)
\end{eqnarray}

It could be seen that, in the log-transformed domain, not only the expected value (mean) of the observable variance is constant, its distribution is also independent of $\sigma$. 
%We have seen this analytically
The above analysis is applicable for two sample variances. 
For larger number of samples $n > 2$, Monte-Carlo simulation is used to visualise the distribution of sample variances. 
The simulation is described as follows:

\begin{enumerate}
\item A large number of input samples are generated ($n \cdot 10^6$ in the figure \ref{fig:variance}) from the PDF given in Eqn. \ref{eqn:log_intensity_base_2_dist}.
\item The input samples are the divided into $10^6$ independent sets, each set consists of $n$ samples.
\item For each input set, the sample variance is calculated.
\item For all input sets, a histogram of the calculated values is then plotted.
\item The consistent variance property is observable when the same plots are obtained, regardless of the value $\sigma$ used in simulations.
\end{enumerate}

\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[simulation match analysis]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_intensity_variance_hist_model_cdf_scene1.eps} 	
		 \label{amplitude}
	} 
	\hfill
	\subfloat[simulation for different number of samples]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_intensity_variance_no_of_samples_cdf.eps} 	
		 \label{intensity}
	}
\end{tabular}
\caption{ Sample Variance: Observable and Modelled CDF Response }
\label{fig:variance}
\end{figure}

Fig. \ref{fig:variance} not only provides evidence for the validity of the simulation process,
  it also visualizes the shapes of the CDF for sample variance distribution of different look numbers.

\subsection{Sample variance as a statistical measure of homogenity}

The discussion in the previous sub-section is based on the assumption of single stochastic processes, and thus homogeneity.  
In a real given analysis area in SAR images, such condition is unknown. 
The principle of SAR speckle filtering however is to group pixels into homogeneous areas, allowing stochastic components to be removed. 
In order to do that, a methodology is required to assert a partitioned area as being homogeneous. 
Fortunately the sampling distribution of log-variance provides a consistent method for testing this homo/hetero-geneity condition.

Under the null hypothesis that the given analysis area is homogeneous,
  the theoretical distribution of sample variance has been given earlier. 
The PDF, and equivalently the CDF, of this distribution can be visualized from Monte-Carlo simulations, as shown in Fig. \ref{fig:variance}. 
Thus, given a computed value of the sample variance in the log-transformed domain,
  the notion that a given analysis area is heterogeneous or homogeneous can be put to test. 
%This forms the basis of the statistical homogeneity hypothesis testing. 
%The test statistic to be used is the sample variance. 
%Figure \ref{fig:variance} shows the simulated CDF for sampling distribution of variance. 
%It is clear that,
Specifically given a pre-determined threshold-value and number of samples,
  if the observable sample variance is higher than a cut-off value, the null hypothesis can be rejected. 
%The collection of cut-off values for a certain threshold-value (a column in table \ref{tab:var_cut_off_values}) form a threshold-vector.

The steps for finding these cut-off values are given below:

\begin{enumerate}
\item For each number of samples, a CDF plot is derived using the Monte-Carlo simulation described earlier.
  \item Intrapolating from the CDF plot, a corresponding cut-off value is derived for each threshold-level.
  \item The combined results form a table, similar in structure to Table \ref{tab:var_cut_off_values}.
  \item The first 3 steps are repeated a few times,
     and the summary statistics (i.e. mean and variance) of the cut-off values for each combination of threshold-value and look-number is reported.  
\end{enumerate}

Table \ref{tab:var_cut_off_values} gives the result of one such such experiment for an arbitrarily chosen set of threshold values.
The table is useful in cases when an area is to be tested for homogeneity,
  for example in the kMLE speckle filter \cite{Le_2010_ACRS}

\begin{table}
\centering

%awk -F"&" '{print $4}' tmp.0.txt | awk -F"(" '{print $1}'
\begin{tabular}{c|c|c|c|c|c}
No  & 70\%            &75\%             &80\%             &85\%             &90\%\\
\hline
2   &2.9024 (0.0252)  &3.7088 (0.0261)  &4.7960 (0.0327)  &6.3488 (0.0242)  &8.7974 (0.0370)\\
3   &3.4698 (0.0141)  &4.1308 (0.0177)  &4.9585 (0.0245)  &6.1176 (0.0136)  &7.9207 (0.0316)\\
4   &3.6505 (0.0225)  &4.2096 (0.0386)  &4.9392 (0.0347)  &5.9195 (0.0341)  &7.4209 (0.0162)\\
5   &3.7421 (0.0134)  &4.2513 (0.0165)  &4.8967 (0.0297)  &5.7729 (0.0170)  &7.0725 (0.0250)\\
6   &3.7850 (0.0149)  &4.2588 (0.0242)  &4.8611 (0.0160)  &5.6449 (0.0162)  &6.8062 (0.0282)\\
7   &3.8274 (0.0119)  &4.2686 (0.0102)  &4.8179 (0.0117)  &5.5447 (0.0114)  &6.6049 (0.0158)\\
8   &3.8468 (0.0110)  &4.2591 (0.0124)  &4.7747 (0.0176)  &5.4519 (0.0087)  &6.4370 (0.0085)\\
9   &3.8610 (0.0081)  &4.2503 (0.0125)  &4.7430 (0.0120)  &5.3688 (0.0085)  &6.2829 (0.0112)\\
10  &3.8663 (0.0089)  &4.2423 (0.0126)  &4.6967 (0.0102)  &5.3015 (0.0103)  &6.1574 (0.0116)\\
11  &3.8707 (0.0089)  &4.2284 (0.0109)  &4.6693 (0.0078)  &5.2387 (0.0111)  &6.0473 (0.0122)\\
12  &3.8763 (0.0063)  &4.2228 (0.0072)  &4.6370 (0.0139)  &5.1805 (0.0050)  &5.9502 (0.0044)\\
13  &3.8808 (0.0048)  &4.2089 (0.0083)  &4.6096 (0.0055)  &5.1315 (0.0092)  &5.8654 (0.0070)\\
14  &3.8761 (0.0048)  &4.1920 (0.0112)  &4.5816 (0.0069)  &5.0845 (0.0058)  &5.7870 (0.0106)\\
15  &3.8806 (0.0048)  &4.1901 (0.0063)  &4.5592 (0.0088)  &5.0372 (0.0076)  &5.7158 (0.0119)\\
16  &3.8774 (0.0032)  &4.1778 (0.0060)  &4.5397 (0.0049)  &4.9985 (0.0087)  &5.6483 (0.0083)\\
17  &3.8760 (0.0043)  &4.1682 (0.0035)  &4.5149 (0.0060)  &4.9652 (0.0051)  &5.5919 (0.0078)\\
18  &3.8708 (0.0083)  &4.1584 (0.0026)  &4.4951 (0.0030)  &4.9273 (0.0043)  &5.5319 (0.0083)\\
19  &3.8694 (0.0055)  &4.1464 (0.0037)  &4.4755 (0.0039)  &4.8961 (0.0027)  &5.4838 (0.0031)\\
20  &3.8687 (0.0051)  &4.1364 (0.0041)  &4.4555 (0.0072)  &4.8673 (0.0042)  &5.4371 (0.0076)\\
21  &3.8672 (0.0064)  &4.1268 (0.0049)  &4.4401 (0.0052)  &4.8360 (0.0045)  &5.3927 (0.0027)\\
22  &3.8637 (0.0046)  &4.1215 (0.0037)  &4.4253 (0.0050)  &4.8136 (0.0043)  &5.3499 (0.0060)\\
23  &3.8610 (0.0032)  &4.1110 (0.0039)  &4.4093 (0.0057)  &4.7870 (0.0059)  &5.3118 (0.0056)\\
24  &3.8571 (0.0033)  &4.1015 (0.0051)  &4.3942 (0.0060)  &4.7650 (0.0026)  &5.2732 (0.0054)\\
25  &3.8541 (0.0054)  &4.0931 (0.0052)  &4.3803 (0.0032)  &4.7416 (0.0024)  &5.2453 (0.0045)
\end{tabular}

\caption{Cut off values}
\label{tab:var_cut_off_values}
\end{table}

%\section{Chapter Conclusion}

%It could be seen that, in the log-transformed domain, not only is the expected mean of the observable variance 
%constant, the sampling distribution of this random variable is also independent of $\sigma$. 
%While we have observed this analytically for variances based on two samples, for a larger number of samples, 
%Monte-Carlo simulation can be used to visualise the PDF of samples' variances.

\textbf{In summary}, a few measurements of distance, being dispersion and contrast, are shown to agree with the consistent sense of variance and homoskedasticity.
The dispersion measurement can be used to test 
if a pixel belongs to a ``known'' class of physical scatterer \cite{Le_2010_ACRS}.
The consistent sense of contrast can be used to test if any pair of measured data points belong to  the same homogeneous class.
An example application of this is described in our fMLE speckle filter \cite{Le_2011_ACRS}.
%It may also explain why, in the original SAR domain, the ratio based 
%  detector / classifier is preferred to differential measures, which are not consistent. 
The consistent sense 
of contrast also gives rise to consistent measures of sample variances, which for example, can be used to test 
if a group of pixels can form a homogenous area.
This was applied in a clustering algorithm described in  \cite{Le_2010_ACRS}.

%\section{Original Heteroskedastic Model}

%\section{The effects of Heteroskedasticity}

%\section{The Homoskedastic effect of Logarithmic Transformation}

%\section{Consistent Measures of Distance in Log-transformed domain}

%\section{Consistent Measures of Variance in Log-Transformed Domain}