### 2013-03-01 5/applications.tex

\chapter{Using the Consistent Measures of Distance} %chapter 5
\label{chap:applications}

\section{Clustering SAR images using the consistent measures}
%\section{SAR Image Clustering using the consistent measures}

%\section{Proposed SAR speckle filter using Homoskedastic features}
\label{sec:k_mle}

Given an area of analysis, 
  the question of whether it is homogeneous or not can be investigated using
%its non-homogeneity can be detected by 
the statistical hypothesis test described above. 
If that is decided to be the case, clustering algorithm can be applied to partition the area into smaller regions. 
Thus an image can be segmented into homogeneous areas by applying the process repeatedly.

%Given an area of analysis, its non-homogeneity can be detected by the statistical hypothesis test described above. 
%If that is likely to be the case, clustering algorithm can be applied to partition the area into smaller regions. 
%The process can be applied iteratively until all partitions can be considered homogeneous (accepting the null hypothesis). 
%This forms the basis of our proposed speckle filtering technique.

\subsection{Developing kMLE clustering Algorithm}

\subsubsection{Homogeneous Maximum Likelihood Estimation}

In a window of analysis, intensity dataset $X^I=\{x^I_1,...,x^I_n\}$ follows the exponential distribution given earlier. 
Assuming prior knowledge of it being homogeneous, thus having a single $\sigma$, the aim then is to estimate $\sigma$. 
This section will show how $\sigma$ can be estimated in log-transformed domain.

It has already been proven that the log-transformed dataset $X^l=\log_2 \sigma$ follows a Fisher-Tippet distribution $pdf(x)= \ln 2 \frac{ 2^d }{ e^{2^d} }$, with $d=x-\mu$, and $\mu=2 \log_2 \sigma$. 
Thus the problem of estimating $\sigma$, casting into the log-transformed domain, equals to the problem of estimating $\mu$. 
The PDF given in Eqn. \ref{eqn:conditional_prob} then can be interpreted as
  the conditional probability of observing a single value of $x^l_i$ when the underlying coefficient equals $\mu$: %is estimated, $\sigma$ can also be estimated $\sigma=2^{ \frac{\mu}{2} }$.

\begin{eqnarray}
\label{eqn:conditional_prob}
P(x^l_i | \mu ) 
	&=& \ln 2 \frac{ 2^{x^l_i - \mu} }{ e^{2^{ x^l_i -\mu }} }
%\displaystyle{ \frac{ \partial }{\partial{ \mu }} P(x^l_i | \mu )  }
%	&=& - \ln^2{2} \frac{ 2^{x^l_i - \mu} (2^{x^l_i - \mu}-1) }{ e^{2^{ x^l_i -\mu }} } 
\end{eqnarray}

Since log-transformed variables are biased, an MLE approach is used to estimate $\mu$. 
The MLE strategy aims to maximize the conditional (likelihood) probability of observing a set of value X found in the area, given the underlying coefficient equals $\mu$: $P(X^l | \mu )$. 
Assuming the values of $x^l_i$ are independent, the statistical product rule indicates that $P(X^l | \mu )=\displaystyle{\prod^n_{i=1}{P(x^l_i | \mu ) }}$, with  $P(x^l_i | \mu)$ as given in equation \ref{eqn:conditional_prob}.   
%$P(x^l_i | \mu ) = \ln 2 \frac{ 2^{x^l_i - \mu} }{ e^{2^{ x^l_i -\mu }} }$. 
Maximize $P(X^e | \mu )$ also maximizes $\ln[P(X^l | \mu )]$, which can be written as:

\begin{eqnarray}
\ln[P(X^l | \mu )] 
	= \displaystyle{\sum^n_{i=1}{ \ln(P(x^l_i | \mu )) }} 
	= \displaystyle{\sum^n_{i=1}{ \ln \left[ \ln 2 \frac{ 2^{x^l_i - \mu} }{ e^{2^{ x^l_i -\mu }} } \right] }} 
%	&=& n \ln \ln 2 +  \displaystyle{\sum^n_{i=1}{ \ln \left( 2^{x^l_i - \mu} \right) - \ln \left( e^{2^{ x^l_i -\mu }} \right) }} 
\end{eqnarray}

The parameter $\mu$ which maximizes the likelihood is found by setting its derivatives to zero:
\begin{eqnarray}
\displaystyle{ \frac{ \partial{ \ln[P( X^l | \mu )] }  }{\partial{ \mu }} }
	&=& 0 \\
\displaystyle{ \frac{ \partial }{\partial{ \mu }} \sum^n_{i=1}{ \ln( 2^{x^l_i - \mu}) - \ln( e^{2^{ x^l_i -\mu }} ) } }
	&=& 0 
%\displaystyle{ \frac{ \partial }{\partial{ \mu }} \sum^n_{i=1}{ \ln( 2^{x^l_i - \mu}) } }
%	&=& - n \cdot \ln2  \\
%\displaystyle{ \frac{ \partial }{\partial{ \mu }} \sum^n_{i=1}{ \ln( e^{2^{ x^l_i -\mu }} )i } }
%	&=& - \ln2 \cdot \sum^n_{i=1}{ 2^{ x^l_i -\mu } } \\
%	&=& - \ln2 \cdot \frac{ \sum^n_{i=1}{ 2^{ x^l_i } }}{2^\mu}  
\end{eqnarray}

Thus $\mu$ can be found by:
\begin{eqnarray}
%n \cdot \ln2	
%	&=& \ln2 \cdot \frac{ \sum^n_{i=1}{ 2^{ x^l_i } }}{2^\mu}  \\
2^\mu 
	&=& \frac{ \displaystyle{ \sum^n_{i=1}{ 2^{ x^l_i } }} }{n}  
\label{eqn:mle_log_transform}
\end{eqnarray}

This result in the log-transformed domain is consistent with the analysis result in the original domain where $\sigma^2 = \frac{\sum^n_{i=1}{x^e_i}}{n} $, with $x^e_i = 2^{x^l_i}$

\subsubsection{Heterogeneous k-MLE Clustering}

In order to apply MLE into a given analysis window ($L^w$), one would need to partition the area into groups of homogeneous pixels ($G$). 
Assuming for now that the number of homogeneous groups ($c=size(C)$) is known \textit{a priori}, this section will describe the proposed clustering algorithm. 
The basic template is similar to the k-means clustering algorithms (Listing \ref{lst:k_mle_algo_template})

\begin{enumerate}
\item Initialize the cluster centres
\item Assign each sample to a centre $C_i$ that has the highest probability $P(x|C_i)$ (as in equation \ref{eqn:conditional_prob})
%(winner take all)
\item Recalculate cluster centres for each $C_i$ (as in equation \ref{eqn:mle_log_transform})
\item Iterate above 2 steps until there is no more changes in cluster centres' locations
\end{enumerate}

%The pseudo code for the algorithm is provided in Listing \ref{lst:k_mle_algo_cluster_centers}

\lstset{language=Matlab,caption=k-MLE Clustering Algorithm Template,label=lst:k_mle_algo_template}
\definecolor{darkgray}{rgb}{0.95,0.95,0.95}
\lstset{backgroundcolor=\color{darkgray}}
\begin{lstlisting}[frame=tb]
function optimized_clusters = k_mle_clustering(l, c_init) 
  c_prev = c_init;
  c_new = find_new_centers(l, c_prev);

  while (c_new != c_prev) do
    c_prev = c_new;
    c_new = find_new_centers(l, c_prev);
  end while

  return optimized_clusters = c_new
end function
\end{lstlisting}

Due to the special characteristics of the log-transformed domain, we have adopted
an MLE estimator in lieu of a standard averaging estimation for cluster centers. 
Consequently the kMLE algorithm is presented in Listing \ref{lst:k_mle_algo_cluster_centers}.
%a standard k-means algorithm transformed into k-mle algorithm given in Listing \ref{lst:k_mle_algo_cluster_centers}.

\lstset{language=Matlab,caption=Iteratively find better cluster centers,label=lst:k_mle_algo_cluster_centers}
\definecolor{darkgray}{rgb}{0.95,0.95,0.95}
\lstset{backgroundcolor=\color{darkgray}}
%\begin{figure}
%\begin{verbatimtab}
\begin{lstlisting}[frame=tb]
function c_new = find_new_centers(l, c)
  for (i = 1 to size_of(l))
    for (j = 1 to size_of(c))
      m[i,j] = 0; %reset memberships
      p[i,j] = conditional_probability_of( l[i]|c[j] );
    end for

    j_0 = argmax( p[i,:] );
    m[j_0] = 1; %assign membership to a cluster
  end for

  for ( j = 1 to size(c) )
    c_new[j] = maximum_likelyhood_of( l[i] with m[i,j]==1 );
  end for

  return c_new;
end function
\end{lstlisting}
%\end{verbatimtab}
%\caption{Iteratively find better cluster centers}
%\label{src:k_mle_algo}
%\end{figure}

When compared with standard k-means the following differences are evident:
\begin{enumerate}
  \item In k-means, data points are grouped into clusters with minimum distance. 
In our specific situation, as the dispersion PDF is non-symmetric around the cluster center, log-transformed data points are assigned to the cluster that gives them highest likelihood probability.
  \item The calculation of new cluster centers is different also. 
In k-means, new cluster centres are estimated as the average of all membered data points. 
Here, due to the bias of the PDF, maximum likelihood estimations are preferred.
  \item All the design decisions are made possible by the knowledge of the statistical PDF in the log-transformed domain!
\end{enumerate}

\subsubsection{Determining number of clusters}

In the previous section the algorithm was designed with the assumption that the number of clusters are known \textit{a priori}. 
In practice, this is not the case. This section will describe how the issue is to be tackled.

The basic idea is to start with an assumption that the whole area is homogeneous.
Then such an assumption is put to test using the hypothesis testing procedure described earlier. 
If the assumption appears to be violated, the area is broken down by introducing a new cluster center. 
The clustering algorithm is again performed and the new clusters are again assumed to be homogeneous, and tested. 
The process iterates until all partitions are confirmed to be homogeneous. 
The pseudo code for such a procedure is given in Listing \ref{lst:k_mle_algo_partition_number}

\lstset{language=Matlab,caption=Homogenous partitions number growing,label=lst:k_mle_algo_partition_number}
\definecolor{darkgray}{rgb}{0.95,0.95,0.95}
\lstset{backgroundcolor=\color{darkgray}}
%\begin{figure}
%\begin{verbatimtab}
\begin{lstlisting}[frame=tb]
function centers = partition_into_homogenous_groups(l, threshold)
  c_init = maximum_likelyhood_of(l);
  [further_partition_needed, c_new] = 
          check_homogeneity_status(l, c_init, threshold);

  while (further_partition_needed) do
    c_prev = c_new;
    c_new = k_mle_cluster(l, c_prev);
    [further_partition_needed, c_new] = 
            check_homogeneity_status(l, c_prev, threshold);    
  end while
  %after this: all paritions are homogenous

  return centers = c_new; 
end function

function [further_partition_needed, c_new] 
              = check_homogeneity_status(l, c, threshold_vector);

  m[i,j] = get_membership_assigments(l, c);
  all_partitions_homogenous = true;
  c_new = c;
  for (j = 1 to size_of(c)) 
    g = {L[i] with m[i,j]==1};
    this_partition_homogenous = 
      (variance(g) < threshold_vector[size_of(g)]);

    if (NOT this_partition_homogenous) then
      % split non-homogenous partition
      c_new[j] = min(g);
      c_new[size_of(c_new) + 1] = max(g);

      all_partitions_homogenous = false;
    end if
  end for

  further_partition_needed = NOT all_partitions_homogenous;
end function
\end{lstlisting}

%\section{Experiment Results and Discussions}%describe analysis and compare
\subsection{Evaluating the kMLE algorithm}
\label{sec:k_mle_result}

\subsubsection{Using the kMLE clustering algorithm for speckle filtering}

It should be noted there is a difference between accepting the null hypothesis and simply failing to reject it. 
However, in the section that follows, the decision rule in use is that if the sample variance is lower than the cut-off value, the null hypothesis is accepted. 
The next analysis shows that for the purpose of speckle-filtering the difference is negligible.

The principle of speckle filtering is to group pixels into homogeneous areas so that the stochastics noise can be removed.
This allows the kMLE clustering algorithm to form the basis of a speckle filtering technique,
  where the results of the clustering algorithm can be transformed into answers for the problem of speckle filter.
Specifically the clustered areas directly serves as the grouping of pixels into homogeneous areas, and the cluster centers serves as the estimated backscattering coefficient for the analysis window.

Let us now consider some extreme threshold-values used as the homogeneous criteria for grouping pixels in the kMLE speckle filter.
If threshold-value is set to a very high value (close to 100\%), then a given partition would almost never be broken down.
The kMLE speckle filter effect would then equals to a box-car filter. 
Conversely, if the threshold-value is set to a very low value (close to 0\%), then a given partition would almost always be broken down to a single value.
The kMLE speckle filtering effect then would be virtually negligible.

Figure \ref{fig:k_mle_diff_threshold} shows various degrees of speckle filtering controllable by using different threshold-values.
Using the same moving window size, here $3\times3$, the effect of filtering is apparently proportional to the threshold values employed:
  with sub-figures \ref{fig:k_mle_threshold_original} to \ref{fig:boxcar_filter} visualise qualitatively the effects of the kMLE speckle filter when the threshold is set at 0\% (or left untouched), 25\%, 80\% and 100\% (or equals to the boxcar filter) respectively.

%	Fig. \ref{fig:k_mle_threshold_original} shows the original intensity image the speckle left untouched (0\%), 
%	Fig. \ref{fig:k_mle_threshold_25pc} and Fig. \ref{fig:k_mle_threshold_80pc} visualise qualitatively the effects of the k-MLE filter when the threshold is set at 25\% and 80\% respectively, 
%	while Fig. \ref{fig:boxcar_filter} shows that too much filtering (boxcar, i.e. 100\%) may lead to blurring effects.

\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[intensity image]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		  \epsffile{images/scene2_intensity.eps}
		 %\epsffile{images/scene2_log_intensity.eps} 	
		 \label{fig:k_mle_threshold_original}
	}
	\hfill
	\subfloat[3x3 25\% k-MLE filter]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/scene2_filtered_LTH_config_04.eps}
		 %\epsffile{images/scene2_log_intensity.eps} 	
		 \label{fig:k_mle_threshold_25pc}
	}  \\
	\subfloat[3x3 80\% k-MLE filter]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/scene2_filtered_LTH_config_03.eps}
		 %\epsffile{images/scene2_filtered_LTH_config_02.eps} 	5x5 k-MLE blur
		 \label{fig:k_mle_threshold_80pc}
	}
	\hfill
	\subfloat[3x3 boxcar filter]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/boxcar_3x3.eps} %images/boxcar_5_blur.eps}
		 \label{fig:boxcar_filter}
	} 
\end{tabular}
\caption{Speckle filtering effects by various threshold-value}
\label{fig:k_mle_diff_threshold}
\end{figure}

One way to improve the speckle suppression effect is by working on a larger window size. 
To reduce the associated blurring effect,
  a lower threshold-value than 100\% compared to the box-car filter would help
    by providing an adaptive partitioning of the analysis window.
Figure \ref{fig:k_mle_iterative_vs_boxcar} shows the visual performance of k-MLE filtering on $5 \times 5$ window.

\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[5x5 boxcar]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		  \epsffile{images/boxcar_5x5.eps}
		 %\epsffile{images/scene2_log_intensity.eps} 	
		 \label{fig:boxcar_5x5}
	} 
	\hfill
	\subfloat[5x5 k-MLE]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/scene2_filtered_LTH_config_02.eps} 	
		 \label{variance}
	} 
\end{tabular}
\caption{Blur effects of larger window size}
\label{fig:k_mle_iterative_vs_boxcar}
\end{figure}

%\subsubsection{Quantitative Evaluation in Log-Transformed domain}

Another way to improve the raw kMLE speckle filter is to apply it iteratively.
As can be seen in figure \ref{fig:k_mle_diff_threshold}, a normal 3x3 k-mle filter would not filter as much noise as a similar boxcar filter would. 
At the same time, a 5x5 boxcar filter while having better speckle suppression power, would lose out by the blurring effects (see Fig. \ref{fig:boxcar_5x5}). 
To improve k-MLE filtering power while avoiding the blurring effect, the kMLE speckle filter can be used iteratively. 
Presented in the figure \ref{fig:k_mle_filtering_power} is the result of a k-MLE filter configured in a 3x3 window with 25\% threshold-value, being applied iteratively for four times. 
%The filtering power is illustrated and the plot shows similar performance with a 5x5 boxcar filter, while the filtered image is rendered with much less blurring effect.
From the plot, the $3 \times 3$ iterative k-MLE filter appears to have similar quantitative performance with the $5 \times 5$ boxcar filter,
  while its visual imagary output appears much sharper, qualititatively speaking.

\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[3x3 25\% 4 iteration k-MLE]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/scene2_filtered_LTH_config_01.eps}
		 \label{variance}
	} 
	\hfill
	\subfloat[Homogenous Error in log-domain]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/k_mle_filter_homogenous_performance.eps}
		 %\epsffile{images/scene2_filtered_LTH_ENL.eps} 
		 \label{variance}
	} 
\end{tabular}
\caption{Improving Homogenous error with iterative k-MLE filter}
\label{fig:k_mle_filtering_power}
\end{figure}

\subsubsection{Evaluation of the kMLE speckle filter on Simulated targets}

A number of point targets were simulated on a homogeneous background clutter. 
The ground truth SNR between the target and background clutter is fixed at 9dB. 
The original, speckled images are presented in Figure \ref{fig:sar_point_target_test_boxcar_k_mle}.
Subsequently, the kMLE speckle filter is to be benchmarked against the standard boxcar filters.
%Both qualitative and quantitative comparisons is done between boxcar filters and the proposed iterative k-MLE filter.
Qualitatively speaking, it is apparent that the kMLE speckle filter delivers
a resultant image with both sharper edges than the 5x5 boxcar filter and a stronger visual contrast  than the 3x3 boxcar filter.
%a sharper image than the 5x5 boxcar filter, at the same time its result has better contrast than the 3x3 boxcar filter.
Quantitatively, the k-MLE also appears to have the best histogram response, of all the 3 filters compared.

\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[5x5 boxcar filter]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		  \epsffile{images/point_target_5x5_boxcar_filtered.eps}
		 \label{intensity}
	}
	\hfill
	\subfloat[iterative k-MLE]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/point_target_iterative_k_mle_filtered.enhanced.eps}
		 \label{variance}
	} \\ 
	\subfloat[3x3 boxcar filter]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		  \epsffile{images/point_target_3x3_boxcar_filtered.eps}
		 \label{intensity}
	}
	\hfill
	\subfloat[histogram response]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/point_target_histogram_preservation.eps}
		 \label{variance}
	} 
\end{tabular}
\caption{Qualitative evaluation on point targets }
\label{fig:sar_point_target_test_boxcar_k_mle}
\end{figure}

\subsubsection{Evaluation of the kMLE speckle filter on Real Images}

The filters are applied on a SLC RadarSat 2 image covering the Muada Merbok area of Malaysia. 
Figure \ref{fig:k_mle_on_RADARSAT2_data} shows the patches of original and filtered images of both natural and urban landscape.
Apparently the kMLE speckle filter delivers a significantly ``better'' image than the original data.

\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[urban: intensity image]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		  \epsffile{images/scene2_intensity.eps}		  	
		 \label{intensity}
	}
	\hfill
	\subfloat[urban: 3x3 iterative k-MLE]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/scene2_filtered_LTH_config_01.eps}
		 \label{variance}
	} \\ 
	\subfloat[natural: intensity image]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		  \epsffile{images/scene_natural_intensity_image.eps}
		 \label{intensity}
	}
	\hfill
	\subfloat[natural: 3x3 iterative k-MLE]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/scene_natural_iterative_k_mle_filtered_image.eps}
		 \label{variance}
	} 
\end{tabular}
\caption{Qualitative evaluation on urban and agricultural areas}
\label{fig:k_mle_on_RADARSAT2_data}
\end{figure}

\section{SAR Speckle filtering using the consistent measures}

\begin{tabular}{c c}
\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:boxcar}
f_{boxcar} = \displaystyle{\sum_{i=1}^n{I_i}} / n 
\end{equation}
\end{minipage}
\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:kmle}
f_{kMLE }= \displaystyle{\sum_{i=1}^k{I_i} } / k 
\end{equation}
\end{minipage}
\end{tabular}

\subsection{The FMLE SAR Speckle Filters}

SAR is stochastic and SAR speckle filtering is the removal of stochastic ``noise''.
The principle of many speckle filters in removing stochastic noise is to group data into homogeneous areas.
In the boxcar filter, the filtering process can be described via Eqn. \ref{eqn:boxcar}, 
where 
	$I_i$ is the intensity value of pixel \textit{i} and
	$n$ is the total number of pixel in the surrounding areas.
The implicit assumption is that ALL pixels in the surrounding area belongs to the same homogeneous area with the estimating central point.
%Clearly, in real-life images, the assumption in boxcar filter is too easy to get violated.
In a filter previously proposed by the authors \citep{Le_2010_ACRS}, the surrounding region is partitioned into homogeneous areas and the approximation is given in Eqn. \ref{eqn:kmle},
where 
	$I_i$ is the intensity value of pixel \textit{i} and
	$k$ is the total number of pixels in the partitioned and homogeneous area covering the centre point.

\begin{tabular}{c c}
\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:boxcar}
f_{boxcar} = \displaystyle{\sum_{i=1}^n{I_i}} / n 
\end{equation}
\end{minipage}
\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:kmle}
f_{kMLE }= \displaystyle{\sum_{i=1}^k{I_i} } / k 
\end{equation}
\end{minipage}
\end{tabular}

By and large, while it appears logical to classify surrounding pixels as being either in the same homogeneous area with the central point or not,
  the concept of ``homogeneous area'' lacks precise definition and boundaries.
This is self-evident in the need for a threshold on the variance measured to assert homogeneity in an ensemble of stochastic samples.

This paper presents a different approach, that of fuzzy logic.
Fuzzy logic embraces truth values that range from 1 to 0, 
	instead of fixing on a rigid two-member set of logical values (i.e. true and false).
Thus instead of making an YES/NO decision on whether or not a data point is in the same homogeneous region with the estimating central point, a fuzzy probabilistic scale is calculated.
Then the noise-removal impact of the surrounding data point will be scaled with this possibilistic probability, in contrast to the usual scheme that proceeds with ``full impact'' or ``no impact at all''.

The idea presented in this paper is structured as follows:
Section \ref{sec:fmle_estimation} will describe the FMLE estimation in details.
The performance of the proposed filter will be evaluated in Section \ref{sec:results_eval}.
Finally, Section \ref{sec:conclusion} concludes this paper.
%The last section will offers some conclusion and finalizes the paper.

The foundation of the approach is the consistent sense of distance found in the log-transformed domain.
It is shown in \cite{Le_2010_ACRS} that given two SAR intensity samples coming from the same background, their distance in the log transformed domain follows a fixed distribution.

Log transformation is defined as:
${L_I}^i = \log_2{(I^i)}$ 
where 
	$I^i$ is the \textit{$i^{th}$} intensity sample in the originally measured domain,
	$L_I^I$ is the value of that sample in the log-transformed domain.
%Distance in log-transformed domain is defined as:
Log domain distance is defined as:
$D= L_I^1 - L_I^2$. 
Assumming $I^i$ follows a negative exponential distribution \citep{Goodman_JOptSocAm_76}, 
  then distance will follow the distribution 
$pdf(D)= \frac{2^D}{(1+2^D)^2} \ln2$,
  which is consistent.
The analysis is confirmed by real-life data validations (see \cite{Le_2010_ACRS}) and visually demonstrated in Fig. \ref{fig:consistent_contrast_log_domain}.

\subsubsection{FMLE Estimation}

Using a fuzzy logic approach, a pixel with intensity value $I_i$ in the surrounding area of the center point $I_0$ has a likelihood of being in the same homogeneous area given as: 
$p_i = pdf(D_i)$, with $D_i=log_2(I_i) - log_2(I_0)$.
We then propose a new approximation given in Eqn. \ref{eqn:fmle_incl}

%We will see how this works in an homogeneous simulated area.
%The figures demonstrate speckle being filtering effects.
In implementating this estimator, we noticed there are visible black dots in the filtered image.
These dots can be explained as the long-tailed nature of the intensity PDF, 
  which manifests itself in the significant existence of very small values in the population.
In the all-inclusive scheme, this will results in 
	a single $D_0=0$ (at $I_0$) and other very large $D_{i, (i > 0)}$, hence
	a single significant $p_0$ (at $I_0$), and other probably insignificant $p_{i, (i>0)}$.
This marginalises the filtering power of surrounding pixels.
An option is to exclude the centre point $I_0$ from its estimation (see Eqn. \ref{eqn:fmle_excl})

\begin{tabular}{c c}
\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:fmle_incl}
FMLE_{incl}= \frac{\displaystyle{\sum_{i=0}^n{p_i * I_i} } }{\displaystyle{\sum_{i=0}^n{p_i } } }
\end{equation}
\end{minipage}
\begin{minipage}[c]{0.45\textwidth}
\begin{equation}
\label{eqn:fmle_excl}
FMLE_{excl}= \frac{\displaystyle{\sum_{i=1}^n{p_i * I_i} } }{\displaystyle{\sum_{i=1}^n{p_i } } } 
\end{equation}
\end{minipage}
\end{tabular}

%\subsection{Comparing the two schemes}
Both filters are applied to an simulated single-look homogeneous scene 
and the histogram of the filtered images are plotted in Fig. \ref{fig:compare_filters_histograms:Original} and Fig. \ref{fig:compare_filters_distance_histograms:Original}.
The filtering power is clearly visible.
%demonstrated in Figure \ref{fig:compare_filters_histograms:Original}. 
In comparing the two schemes' histogram in the same figure, it appears that 
	the filtering power of ignoring centre point scheme is at least as powerful as that of centre point included scheme and
	the skewness in the histogram of the ignoring scheme is also appearing less than that of the included scheme.

\begin{figure}[H]
\centering
\subfloat[Intensity histogram]{
  \epsfxsize=6cm
  \epsfysize=6cm
  \epsffile{images/consistent_contrast_log_domain.made.eps}
\label{fig:consistent_contrast_log_domain}
  }
\caption{Consistent Distance Property and the FMLE Estimations}
\end{figure}

\begin{figure}[H]
\centering
\begin{tabular}{c}
\subfloat[Intensity histogram]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/compare_filters_histograms.3.made.eps} 
\label{fig:compare_filters_histograms:Original}
	}
        \hfill 
	\subfloat[Distance histogram]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/compare_filters_distance_histograms.3.made.eps} 
\label{fig:compare_filters_distance_histograms:Original}
	}\\ 
 	\subfloat[FMLE (incl) outputs\label{fig:compare_filters_distance_histograms:Filtered:incl}]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/consistency_preservation.d_incl.made.eps} 
	} \hfill
	\subfloat[FMLE (excl) outputs\label{fig:compare_filters_distance_histograms:Filtered:excl}]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/consistency_preservation.d_excl.made.eps} 
	} 
\end{tabular}
\caption{Consistent Distance Property and the FMLE Estimations}
\label{fig:Consistent_Distance_Property_and_the_FMLE_Estimations}
\end{figure}

\subsubsection{Preservation of Distance Histogram Consistency}

Fig. \ref{fig:compare_filters_distance_histograms:Filtered:incl} and \ref{fig:compare_filters_distance_histograms:Filtered:excl} plot the distance histograms found in the outputs of applying FMLE filters on the inputs depicted in Fig. \ref{fig:compare_filters_histograms:Original}.
They clearly demonstrate that the filtering treatment described above preserves the consistent sense of distance.
This can be explained from the consistent histogram of the fuzzy possbilistic probabilities.
Unfortunately, the filtering formula given in Eqn. \ref{eqn:fmle_incl} and Eqn. \ref{eqn:fmle_excl}, especially when applied to large image areas, involves large number of random variables.
This renders the task of giving analytical PDF impractical.
However, large scale computer simulations can provide a practical histogram, from which the PDF can estimated.
%from large scale simulations. 
%Simulating experiments with $10^6$ samples is carried out and the histogram information is saved.
Given a distance $d$, its corresponding probability $pdf(d)$ can be simulated by intrapolating from the saved histogram $y\equiv hist(x)$ at the value of $x \equiv d$.

\subsubsection{Recursively Applying Fuzzy MLE Filter}

The preservation of consistent distance property is the enabling factor that allows the recursive use of FMLE filters.
Additional applications of the filtering process can be expected to reduce the random noise variance even further.
In a sense, this scheme is similar to statistical asymptotic estimations, with the added advantage of spatial preservation.

In the subsequent application of the filtering algorithm, the main different when compared with the filters in initial iteration is the way the new fuzzy probabilities being calculated.
This subsequent computation of fuzzy probabilities is based on the derived distance PDF, being computed in the previous section.
This process is iterated iteratively, with the histogram obtained from the previous iteration serves as probabilistic distance PDF input of the subsequent application of the filter.
This scheme allows speckle to be further depressed, as illustrated in Fig. \ref{fig:plot_incl_distance_pdf_by_iterations}

\begin{figure}[h!]
\centering
	\subfloat[FMLE (incl): 123]{
		 \epsfxsize=3.5cm
		 \epsfysize=3.5cm
		 \epsffile{images/incl_distance_pdf_by_iterations.123.made.eps} 
		 \label{fig:plot_incl_distance_pdf_by_iterations.123}
	}
	\subfloat[FMLE (incl): 456]{
		 \epsfxsize=3.5cm
		 \epsfysize=3.5cm
		 \epsffile{images/incl_distance_pdf_by_iterations.345.made.eps} 
		 \label{fig:plot_incl_distance_pdf_by_iterations.345}
	} 
	\subfloat[FMLE (excl): 123]{
		 \epsfxsize=3.5cm
		 \epsfysize=3.5cm
		 \epsffile{images/excl_distance_pdf_by_iterations.123.made.eps} 
		 \label{fig:plot_excl_distance_pdf_by_iterations.123}
	}
	\subfloat[FMLE (excl): 456]{          
		 \epsfxsize=3.5cm
		 \epsfysize=3.5cm
		 \epsffile{images/excl_distance_pdf_by_iterations.345.made.eps} 
		 \label{fig:plot_excl_distance_pdf_by_iterations.345}
	}
\caption{Iteratively Improving in Distance Histograms}
\label{fig:plot_incl_distance_pdf_by_iterations}
\end{figure}

The log-transformed domain provides a consistent sense of distance.
By embracing a probabilistic approach, the proposed Fuzzy Maximum Likelihood Estimator makes use of this consistency.
Experimental results suggest that the estimator's output also exhibits this consistency property, which allows recursive application of the filter discussed above.
The performance of the filter is evaluated both qualitatively against real images and quantitatively via simulated experiments.
The speckle suppression power is evaluated by measuring the ubiquitous MSE on simulated and perfectly homogeneous truth-grounded experiments.
The results suggest that by applying the estimation recursively, this value can be reduced arbitrarily low, assuming that the number of iterations, and hence the available scene and computational power, is sufficiently large.
For heterogeneous scenes, the robust Area Under the ROC Curve criteria is used as statistical evidence for better performance in the subsequent tasks of target detection and surface classification.
Experiments confirm the power of FMLE filters in various heterogeneously simulated areas.
They also exhibit good correlation between the MSE and the AUC index.

\subsection{Evaluating the FMLE SAR speckle filters}

Estimators are evaluated based on the bias and variance properties of their estimates. 
The performance evaluation is usually based qualitatively on some real data and quantitatively through simulated experiments. 
Due to the stochastic nature of simulation process, results of repeated experiments are normally preferred.

The nature of SAR speckle is that
	it is stochastic even if the underlying radiometry is constant. 
And when there are spatial radiometric variation, not only its expected value changes, its heteroskedastic variance varies as well.
Nevertheless, statistical models have been developed to derive the underlying back-scattering coefficient ($\sigma$) from measured SAR data. 
As such, speckle filters are, by and large, estimators that attempt to determine this unknown coefficient from observable SAR data.

SAR speckle-filtering can be, and has been, positioned within the context of estimation theory\cite{Touzi_2002_TGRS}. 
The stages in this statistical framework consist of statistical modeling, estimator development and evaluation of the estimators' performance. 
%Estimators are typically evaluated based on the bias and variance properties of their estimates: 
%	with lower bias and / or lower variance, hence lower MSE indicates better accuracy.
Estimators are typically evaluated based on the bias and variance properties of their estimates:
  lower bias and/or lower variance would lead to lower MSE, which then indicates better accuracy.
Ideally, estimator evaluation should be based both qualitatively on some real data and quantitatively through simulated experiments. 
In addition, due to the stochastic nature of SAR-processing, and thus its simulating process, 
	statistical summaries of repeatedly-simulated experiments are normally preferred to single-run results.

Our survey of SAR speckle filter research, however, indicates 
	a different picture from the standard practice of the statistical framework, 
	specifically, in the stage of performance evaluation. 
Due to the multiplicative and heteroskedastic nature of speckle noise, 
	bias and variance evaluation may not be the most useful measures. 
Thus, the standard evaluation metric of mean-squared-error is not easily applicable,
	prompting alternative evaluation criteria to be proposed.
In fact, our survey of relevant literature fails to reveal 
	a single universally agreed quantitative metric for the performance of speckle filters.

Most commonly, each newly proposed speckle filter tends to be published along with its own methodology for evaluating its 
	performance. 
As such, many papers lack a comparative basis beyond simple visual qualitative comparison on a few image scenes. 
While such visual comparison is useful, as an evaluative methodology, regrettably it lacks scientific objectivity. 
Some papers do present quantitative measurements. 
However, due to the lack of a standardized performance criteria, 
	the evaluation metrics can change significantly from one paper to the next.

The performance of speckle filters are normally evaluated based on a few criteria \citep{Nyoungui_2002_IntlJRemoteSense}.
The most widely used criteria is speckle suppression, 
  which is often evaluated over homogeneous area.
The other important criteria include radiometric preservation, something undeniably desirable in the filtered output.
In practical terms, the most common usage of speckle filtered imagery include target and feature detection and classification.
We highlight the use of the ROC curve and the area under it as the statistical evidence of separability between target and clutter PDFs.
We then show that lower MSE achievable by FMLE filters does lead to better performance in the subsequent tasks of target detection and classification.
%As is customary, we start by an visual evaluation of the filter on real-life images.
We start with visual evaluation of the filters on real-life images.

\subsubsection{Qualititative Evaluation on Real Images}

The filters are applied to a SLC RadarSat 2 image covering the Muda Merbok area of Malaysia.
Fig. \ref{fig:fmle.real.images} shows the patches of original and filtered images of both natural and urban landscape.
The filtering effects are clearly visible.
Apparently the second iteration outputs appears much better than the initial iteration, and offer roughly the same level of performance in comparision to that of boxcar filter.

\begin{figure}[h!]
\centering  
      \begin{tabular}[h]{c}
	\subfloat[Original]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.natural.incl.iter0.made.eps} 
		 \label{fig:fmle.natural.incl.iter0}
	} \hfill
	\subfloat[Boxcar]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/boxcar.natural.3x3.made.eps} 
		 \label{fig:boxcar.natural}
	}\\
	\subfloat[FMLE incl, iter 1]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.natural.incl.iter1.made.eps} 
		 \label{fig:fmle.natural.incl.iter1}
	} \hfill
	\subfloat[FMLE incl, iter 2]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.natural.incl.iter2.made.eps} 
		 \label{fig:fmle.natural.incl.iter2}
	} \\
	\subfloat[FMLE excl, iter 1]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.natural.excl.iter1.made.eps} 
		 \label{fig:fmle.natural.excl.iter1}
	} \hfill
	\subfloat[FMLE excl, iter 2]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.natural.excl.iter2.made.eps} 
		 \label{fig:fmle.natural.excl.iter2}
	} 
	\end{tabular}
\caption{Qualitative Evaluation of FMLE filter on Real Life Scene of Natural Surface}
\label{fig:fmle.real.images}
\end{figure}

\begin{figure}[h!]
\centering  
      \begin{tabular}{c}
	\subfloat[Original]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.urban.excl.iter0.made.eps} 
		 \label{fig:fmle.urban.excl.iter0}
	} \hfill
	\subfloat[Boxcar]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/boxcar.urban.3x3.made.eps} 
		 \label{fig:boxcar.urban}
	} \\
	\subfloat[FMLE incl, iter 1]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.urban.incl.iter1.made.eps} 
		 \label{fig:fmle.urban.incl.iter1}
	} \hfill
	\subfloat[FMLE incl, iter 2]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.urban.incl.iter2.made.eps} 
		 \label{fig:fmle.urban.incl.iter2}
	} \\ 
	\subfloat[FMLE excl, iter 1]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.urban.excl.iter1.made.eps} 
		 \label{fig:fmle.urban.excl.iter1}
	} \hfill
	\subfloat[FMLE excl, iter 2]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/fmle.urban.excl.iter2.made.eps} 
		 \label{fig:fmle.urban.excl.iter2}
	} 
	\end{tabular}
\caption{Qualitative Evaluation of FMLE filter on Real Life Scene of Urban Landscape}
\label{fig:fmle.real.images}
\end{figure}

\subsubsection{Evaluating Speckle Suppression Effects}

We have found that MSE in the log-transformed domain is related to the standard Equivalent Number of Look value.
Deriving from the results of \cite{Hoekman_1991_TGRS} and \cite{Xie_TGRS_2002}, the formula is given as:
$MSE = \frac{1}{(ENL - 0.5)ln^2(2)}$.
%An experiment is carried out to 
Fig. \ref{fig:verify_MSE_as_perf_index} demonstrates the relationship.
The use of iterative Fuzzy MLE estimations allows further suppression of speckle.
This is illustrated with ``tighter'' distance histograms and reducing MSE, as plotted in Fig. \ref{fig:Homogeneous_Area}

\begin{figure}[h]
\centering
 \epsfxsize=6cm
 \epsfysize=6cm
 \epsffile{images/verify_MSE_as_perf_index.made.eps} 
\end{figure}

\begin{figure}[h]
\centering
      \begin{tabular}[h]{c}
	\subfloat[MSE (incl, 3x3)]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/distance_pdf_simulation_iterative.incl.made.eps} 
		 \label{fig:distance_pdf_simulation_iterative.incl}  
	} \hfill
	\subfloat[MSE (excl, 3x3)]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/distance_pdf_simulation_iterative.excl.made.eps} 
		 \label{fig:distance_pdf_simulation_iterative.excl}
	} \\
	\subfloat[MSE (incl, 5x5)]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/distance_pdf_simulation_iterative.5x5.incl.made.eps} 
		 \label{fig:distance_pdf_simulation_iterative.incl.5x5}
	} \hfill
	\subfloat[MSE (excl, 5x5)]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/distance_pdf_simulation_iterative.5x5.excl.made.eps} 
		 \label{fig:distance_pdf_simulation_iterative.excl.5x5}
	}
     \end{tabular}
\caption{Homogeneous Area: MSE criteria and speckle suppression power of FMLE filters}
\label{fig:Homogeneous_Area}
\end{figure}

\subsubsection{Evaluating Target Detection Performance}

The normal application after applying filters on SAR images is to detect the existence of certain target within its surrounding clutters.
%Assume the 
The most common type of detector (or classifier) employs a threshold based approach in determing whether an abnormal value signifies the existence of a target.
The Receiver Characteristic Curve (ROC) and the area under it (AUC) is typically used to evaluate the detectability of target features \citep{Mazurowski_2009_IJCNN}.

It is normally claimed that applying speckle filters increases the performance of subsequent target detection tasks.
The following experiment illustrates this point.
In this experiment, we apply a simple 3x3 boxcar filter to two different homogeneous and SLC noise corrupted scenes that are known to be 3dB apart.
And then the histograms as well as the resulting ROC curve are plotted between the pairs of target and background histogram in the two cases of unfiltered and filtered data.
We noted that the ROC curve for the histograms in original domain matches perfectly with ROC for the histograms in the log-transformed domain (see Fig. \ref{fig:PDF_Discriminality}), as expected from theory.

\begin{figure}[h]
\centering
      \begin{tabular}[h]{c}
 	\subfloat[SLC Intensity]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/illustrate_ROC_curve.pdf.s.made.eps} 
		 \label{fig:illustrate_ROC_curve.pdf.s.made.eps}
	} \hfill
	\subfloat[9-Look Intensity]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/illustrate_ROC_curve.pdf.m.made.eps} 
		 \label{fig:illustrate_ROC_curve.pdf.m.made.eps}
	} \\
 	\subfloat[SLC log-intensity]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/log_illustrate_ROC_curve.pdf.s.made.eps} 
		 \label{fig:log_illustrate_ROC_curve.pdf.s.made.eps}
	} \hfill
	\subfloat[9-Look log-intensity]{
		 \epsfxsize=6cm
		 \epsfysize=6cm
		 \epsffile{images/log_illustrate_ROC_curve.pdf.m.made.eps} 
		 \label{fig:log_illustrate_ROC_curve.pdf.m.made.eps}
	}
   \end{tabular}
\caption{Histogram Discriminality and the ROC Curve}
\label{fig:PDF_Discriminality}
\end{figure}

%\subfloat[Invariant ROC Curve]{
\begin{figure}
 \epsfxsize=6cm
 \epsfysize=6cm
     		 \epsffile{images/log_illustrate_ROC_curve.made.eps} 
		 \label{fig:illustrate_ROC_curve.roc}
  \caption{Invariant ROC Curve}
\end{figure}

The most common types of target to be detected in image processing are point targets, line targets, and edge targets.
The detectability of the targets are measured using the AUC index which is tabulated against the MSE.
Fig. \ref{fig:Heterogeneous_Patterns} illustrates the evaluation result in benchmarking the performance of FMLE filters for heterogenous regions.
Even though the relationship is probably not linear,
  the correlation between MSE and Area Under the ROC curve (AUC) is apparent,
Also FMLE filters apparently have better performance than boxcar filters.

\begin{figure}[h!]
\begin{minipage}[c]{0.5\textwidth}
%\captionsetup[subfigure]{labelformat=empty}
%\subfloat[]{%
%\captionsetup[subfigure]{labelformat=parens}
%\begin{figure}
%\centering
	\subfloat[Point Target]{
		 \epsfxsize=3cm
		 \epsfysize=3cm
		 \epsffile{images/641bcaad-f54b-4bb6-8d10-77784da522a5.made.eps} 
		 \label{fig:point_target_pattern}
	}
	\subfloat[Edge]{
		 \epsfxsize=3cm
		 \epsfysize=3cm
		 \epsffile{images/c5124949-1228-4749-a99e-caf9430e400a.made.eps} 
		 \label{fig:edge_pattern}
	}\\
	\subfloat[Line]{
		 \epsfxsize=3cm
		 \epsfysize=3cm
		 \epsffile{images/a2e330a8-3deb-4328-94d8-d18c43e4f7a1.made.eps} 
		 \label{fig:line_pattern}
	}
	\subfloat[Dotted Pattern]{
		 \epsfxsize=3cm
		 \epsfysize=3cm
		 \epsffile{images/38e79d4c-af61-4b1f-acc2-528ecedb96a7.made.eps} 
		 \label{fig:dotted_pattern}
	}
%\caption{Simulated Heterogeneous Patterns}
%\label{fig:distance_pdf_simulation_iterative}
%\end{figure}
\end{minipage}
%} & 
\begin{minipage}[c]{0.38\textwidth}
%\captionsetup[subfigure]{labelformat=empty}
%\subfloat[]{%
%\captionsetup[subfigure]{labelformat=parens}
%\begin{table}
%\centering
\begin{tabular}{|l|l|l|l|}
\hline
Scene 				& Filter 						& MSE 	& AUC \\
\hline
Point Target 	& FMLE (incl, 3x3) 	& 1.38	& 0.901 \\
Point Target 	& FMLE (excl, 3x3) 	& 1.47	& 0.893 \\
Point Target 	& boxcar (3x3) 			& 1.56	& 0.826 \\
Point Target 	& boxcar (5x5) 			& 2.67	& 0.502 \\
\hline
Edge 	& FMLE (incl, 3x3) 	& 0.68	& 0.969 \\
Edge 	& FMLE (excl, 3x3) 	& 0.73	& 0.964 \\
Edge 	& boxcar (3x3) 			& 1.03	& 0.942 \\
Edge 	& boxcar (5x5) 			& 1.51	& 0.934 \\
\hline
Line 	& FMLE (incl, 3x3) 	& 1.89	& 0.791 \\
Line 	& FMLE (excl, 3x3) 	& 2.02	& 0.767 \\
Line 	& boxcar (3x3) 			& 2.27	& 0.658 \\
Line 	& boxcar (5x5) 			& 2.88	& 0.375 \\
\hline
Dotted 			& FMLE (incl, 3x3) 	& 2.28	& 0.639 \\
Dotted 			& FMLE (excl, 3x3) 	& 2.32	& 0.621 \\
Dotted 			& boxcar (3x3) 			& 2.92	& 0.431 \\
Dotted 			& boxcar (5x5) 			& 3.35	& 0.382 \\
\hline
\end{tabular}
%\caption{The MSE \& ROC indices are correlated}
%\end{table}
\end{minipage}
%}
%\end{tabular}
\caption{Heterogeneous Patterns: MSE criteria and FMLE performance}
\label{fig:Heterogeneous_Patterns}
\end{figure}

\subsubsection{Contribution and Conclusions}

The nature of SAR speckle is stochastic. 
The speckle filtering problem has been casted into a statistical estimation theory framework. 
Speckle filtering, by and large, means the removal of stochastic component in SAR images. 
Various negative effects of heteroskedasticity are discussed. 
Log-transformation is shown to provide homoskedasticity in statistical models. 
Using speckle filtering as an applicative example, these homoskedastic features are shown to be useful. %here to improve speckle-filtering.

The speckle filtering technique proposed here has been shown to work for single-look SAR images. 
Multi-look processed image speckle filtering should be possible to use this technique with some minor adaptations. 
The filter is developed from an easy to understand winner-take-all clustering approach. %to a more fuzzy probabilistic clustering algorithm. 
Should the fuzziness be taken to a new level, an optimal fuzzy probabilistic filter which requires no clustering may be specified and designed.

\section{Evaluating SAR Speckle Filters using the consistent measures}

To do quantitative evaluation of different speckle filters, it is necessary to compare a filtered image with a ground-truth one. 
Furthermore, it is desirable that the error, i.e. the distance between the filtered output and the original input, should be consistent and independent to the particular simulated scenario. 
Normally, the measure of choice is in the original SAR intensity or amplitude domain. 
However owing to the homoskedastic effects of the log-transformation, the error in log-transformed domain is preferred.

Homoskedasticity makes ordinary least square (OLS) approaches becomes valuable again. 
Synonymous to OLS is the use of standard Mean Squared Error as a performance index.

\begin{equation}
\Psi = \displaystyle{ \frac{ \sum^N_{i=1}{(Y^L_i - X^L_i)^2} }{N} }
\end{equation}
where $Y^L_i$ indicates filtered output values and $X^L_i$ denotes the original input as ground-truth values, $N$ is the number of data points available

This section will shows that in a homogeneous area, this performance index is similar to the standard Equivalent Number of Looks (ENL) measures, while investigating the value in heterogeneous areas shows how well the filter adapts to underlying spatial variations.

%\section{Using MSE to evaluate SAR speckle filters}

The nature of SAR speckle is that
	it is stochastic even if the underlying radiometry is constant. 
And when there are spatial radiometric variation, not only its expected value changes, its heteroskedastic variance varies as well.
Nevertheless, statistical models have been developed to derive the underlying back-scattering coefficient ($\sigma$) from measured SAR data. 
As such, speckle filters are, by and large, estimators that attempt to determine this unknown coefficient from observable SAR data.

SAR speckle-filtering can be, and has been, positioned within the context of estimation theory\cite{Touzi_2002_TGRS}. 
The stages in this statistical framework consist of statistical modeling, estimator development and evaluation of the estimators' performance. 
%Estimators are typically evaluated based on the bias and variance properties of their estimates: 
%	with lower bias and / or lower variance, hence lower MSE indicates better accuracy.
Estimators are typically evaluated based on the bias and variance properties of their estimates:
  lower bias and/or lower variance would lead to lower MSE, which then indicates better accuracy.
Ideally, estimator evaluation should be based both qualitatively on some real data and quantitatively through simulated experiments. 
In addition, due to the stochastic nature of SAR-processing, and thus its simulating process, 
	statistical summaries of repeatedly-simulated experiments are normally preferred to single-run results.

Our survey of SAR speckle filter research, however, indicates 
	a different picture from the standard practice of the statistical framework, 
	specifically, in the stage of performance evaluation. 
Due to the multiplicative and heteroskedastic nature of speckle noise, 
	bias and variance evaluation may not be the most useful measures. 
Thus, the standard evaluation metric of mean-squared-error is not easily applicable,
	prompting alternative evaluation criteria to be proposed.
In fact, our survey of relevant literature fails to reveal 
	a single universally agreed quantitative metric for the performance of speckle filters.

Most commonly, each newly proposed speckle filter tends to be published along with its own methodology for evaluating its 
	performance. 
As such, many papers lack a comparative basis beyond simple visual qualitative comparison on a few image scenes. 
While such visual comparison is useful, as an evaluative methodology, regrettably it lacks scientific objectivity. 
Some papers do present quantitative measurements. 
However, due to the lack of a standardized performance criteria, 
	the evaluation metrics can change significantly from one paper to the next.

\subsection{ Speckle Filtering Process And The Homoskedastic Log-transformed Domain}
\label{sec:schema_log_images}

\begin{figure}[h]
 \centering
 \epsfxsize=2.4in
 \epsfysize=1.2in
 \epsffile{images/simulation_schema.eps} 	
\caption{SAR simulation, processing and filtering schema}
\label{fig:simul_process_filter_schema}
\end{figure}

Fig. \ref{fig:simul_process_filter_schema} illustrates the general schema of SAR simulation / processing and 
SAR speckle filtering process.
The ``SAR processor'' block indicates that either one of the following processes can occur.
First is the normal process of SAR processing where the unknown ground radiometric attribute ($X$) is recorded 
in $\tilde{X}$. Second is the simulation process where the noisy SAR data $\tilde{X}$ is simulated from a known 
ground-truth pattern $X$.
%In this paper, fully developed speckle assumption is employed, which allow for easier simulation process.	
%Eventhough different levels of noise can be simulated, in the experiments presented here, only single-look are simulated, and the similar boxcar filter is studied.

The ``SAR Speckle Filter'' block indicates the speckle filtering process, 
	which takes the noisy speckled SAR data $\tilde{X}$ as input
	and outputs $\hat{X}$ as a better estimation of the ground-truth (i.e. $X$).
The speckle filters used in this paper are: 
	boxcar filter, enhanced Lee filter, enhanced Kuan filter, enhanced Frost filter, Gamma Map filter 
	and PDE filter \cite{You_TIP_2000}.
Evidently for the special case where there is no filter applied then $\hat{X}^{none} = \tilde{X}$.
%Thus $MSE_{residual}^{none} = MSE_{base}$

Logarithmic transformation offers several consistent measures of distance, which is hypothised to be 
significant in evaluating speckle filters.
	A case in point is in visual evaluation of filtered results obtained from real captured SAR images.
Since the radiometric ground-truth is not available, %the most common 
%In real SAR captured images, the radiometric ground-truth normally is not available.
	the most common way to evaluate speckle filters then is by qualitative visual evaluation.
A conventional method is to investigate the ratio images between the filtered output and the noisy input images.
Since logarithmic transformation converts these ratio into substractive residual which are consistent, residual 
analysis can be used to analyse and evaluate the performance of speckle filters.
%To compare the use of ratio images in the original domain and the residual image in the log-transformed domain, 
For visual comparison, Fig. \ref{fig:real_image_ratio_vs_residual} depicts the residual image in the log-transformed domain 
and the ratio images in the original domain, where a simple boxcar filter has been applied to data obtained from a 
real RADARSAT SAR image.
``Visible'' structure appears to be more easily discernable in the log-2 residual random pictures than in 
the ratio images, although this conclusion is of course itself a subjective one.

%Our qualititative conclusion is that
%	a. it is apparently easier to notice the ``visible structure'' being removed in the residual noise pictures than the ratio mages, and
%	b. even so, the above conclusion is at best quite subjective.
%Thus to compare the filters against each other, 
%	quantitative measures are preferred.

\begin{figure}
\centering  
\begin{tabular}{c}
	\subfloat[Original Patch]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_real.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Boxcar Filtered Result]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_real.boxcar.eps} 	
		 \label{intensity}
	} \\
	\subfloat[Ratio: Filtered / Original]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_real.ratio2.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Ratio: Original / Filtered]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_real.ratio1.eps} 	
		 \label{intensity}
	}  \\
	\subfloat[Log Residual: Filtered - Original]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_real.residual2.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Log Residual: Original - Filtered]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_real.residual1.eps} 	
		 \label{intensity}
	} 
\end{tabular}
\caption{Visualising Removed ``Noise'': Ratio images in the original domain vs. Residual images in the log-transformed domain}
\label{fig:real_image_ratio_vs_residual}
\end{figure}

\begin{subequations} \label{eqn:eval_metric}
\begin{align}
MSE_{true} = E \left[ (\hat{X}^L - X^L)^2 \right] \\
MSE_{base} = E \left[ (\tilde{X}^L - X^L)^2 \right] \\
MSE_{noise} = MSE_{residual} = E \left[ (\hat{X}^L - \tilde{X}^L)^2 \right] \\
MSE_{benchmark} = \left| MSE_{residual} - MSE_{base} \right| 
\end{align}
\end{subequations}
%\begin{eqnarray}
%MSE_{true} = E \left[ (\hat{X} - X)^2 \right] \\
%MSE_{noise} = E \left[ (\hat{X} - \tilde{X})^2 \right] \\
%MSE_{base} = E \left[ (\tilde{X} - X)^2 \right] \\
%MSE_{benchmark} = \left| MSE_{noise} - MSE_{base} \right| 
%\end{eqnarray}

As qualititative visual evaluation is subjective by nature,
	quantitative metrics and measurements are preferred, if relevant and possible.
Different metrics to evaluate speckle filters, which will be investigated in subsequent sections, 
are given in Eqns. \ref{eqn:eval_metric}, all in log-transformed domain.
When ground-truth is available, 
	either in simulated experiments or over homogeneous area where it can be reasonably estimated, 
	true MSE is the expected squared error between the estimated and true values.

However, in real captured SAR images where such ground truth is unknown, the evaluation can then be carried out 
through a benchmarked MSE. 
The idea is that since both the speckled values and the estimated values are available, the residual $MSE_{noise}$ 
can be computed, which can be thought of as the level of noise being removed by speckle filtering process.
Even as the ground-truth is not known in real SAR images, their speckle level (i.e. ENL) may be known or can be 
estimated reasonably well.
Thus the base MSE level, which is also a measure of the speckle noise level, can also be estimated.
Naturally, the speckle filter should remove as much noise as possible, 
	while it also should not remove more variations than that caused by the speckle noise.
%Assumming that the speckle level of the scene is known (e.g. ENL can be reasonably estimated) then the base

\subsection{Evaluating SAR Speckle Filters Over Homogeneous Areas}

%\subsubsection{Evaluating Speckle Filters on Homogeneous Areas}
%\label{sec:eval_homo}

In this section, a methodology to evaluate speckle filters over homogeneous area in the log-transformed domain is 
illustrated. Under the condition of homogeneity, speckle filters are supposed to estimate with negligible bias.
Then the filters are traditionally compared by measuring their speckle suppression power using the ENL index.
Consequently, in the log-transformed domain, the first component of MSE evaluation (i.e. bias evaluation) is 
probably insignificant in comparison to the other components, namely, variance evaluation. 
Subsection \ref{sec:homogeneous_theoretical} gives a theoretical justification for our methodology where variance 
methods of evaluation are shown to be mathematically related to the ENL index.
The final subsection details how variance evaluation in the log-transformed domain to evaluate the levels of speckle 
over homogeneous areas.

\subsubsection{ Estimating ENL from MSE index in homogeneous areas }
\label{sec:homogeneous_theoretical}

In this subsection, we show that the variance in the log-transformed domain can be related mathematically to the ENL 
index. Over homogeneous area, the ground-truth is unchanged, i.e. $X^L_i=X^L \forall i$.
Assuming the filters achieve negligible bias, i.e. $E(\hat{X}^L)=X^L$, 
	then the MSE evaluation is reducced to variance evaluation.
That is, for known homogeneous scenes, MSE can be estimated as the observable variance of the filtered 
output in the log-transformed domain.

Let us consider the speckle suppression effect of multi-look processing in the log-transformed domain. 
%Since multi-look processing is unbiased, we will take the variance of log-transformed multi-look processing output as the MSE performance index. 
Hoek \cite{Hoekman_1991_TGRS} and Xie \cite{Xie_2002_TGRS} have given the variance for L-look log-transformed random 
variables as: 
\begin{equation}
var(\hat{X}^L)= \frac{1}{\ln^2(2)} \left( \frac{\pi^2}{6} - \sum^{L-1}_{i=1}{\frac{1}{i^2}} \right).
\label{eqn:perf_index_theoretical}
\end{equation}

%Next we will be showing that the ENL, i.e. L, can be estimated from a given $var(Y^L)$. 
Taking results of the Euler proof for the Basel problem, we have $\frac{\pi^2}{6} = \sum^{\infty}_{i=1}{ \frac{1}{i^2} } $, then $var(Y^L)= \frac{1}{\ln^2(2)} \left( \sum^{\infty}_{i=L}{ \frac{1}{i^2} } \right) $.
Noting that $ \frac{1}{i} - \frac{1}{i+1} = \frac{1}{i(i+1)} < \frac{1}{i^2} < \frac{1}{i(i-1)} = \frac{1}{i-1} - \frac{1}{i}$, then $ \frac{1}{L} - \frac{1}{\infty} < \sum^{\infty}_{i=L}{ \frac{1}{i^2} }  < \frac{1}{L-1} - \frac{1}{\infty} $.
In fact, we could estimate:
\begin{equation}
  var(\hat{X}^L) = \frac{1}{(L-0.5) \ln^2(2) }
\label{eqn:perf_index_analytic}
\end{equation}
Thus the ENL, i.e. $L$, can be estimated as:
\begin{equation}
\hat{L}_{est} = \frac{1}{var(\hat{X}^L) \ln^2(2)} + 0.5
\label{eqn:enl_analytic}
\end{equation}

The above analysis can be verified experimentally. First, a $512\times512$ homogeneous area is generated and 
corrupted with single look SAR-speckle PDF. 
Different multi-look processing filters are then applied to the image patch. 
The observable variance in the log-transformed domain is recorded, and 
	 ENL is then estimated from Eqn. \ref{eqn:enl_analytic} for each simulation.
The simulation is performed multiple times, with Table \ref{tab:enl_in_log_domain} reporting the results in terms 
of mean and standard deviation.
The theoretical variances are calculated from Equation \ref{eqn:perf_index_theoretical}, 
	together with the analytical values from Eqn. \ref{eqn:perf_index_analytic}.
These experimental results evidently validate the analysis given above.
They also shows that 
	while the approximation in Eqn. \ref{eqn:enl_analytic} may not be perfect, 
	since ENL is supposed to be an integer, the value obtained actually corresponds very closely with the 
	nearest correct integer result.

\begin{table}[h]
\centering
\begin{tabular}{r|c|c|c|c}
ENL & Var (Theory) & Var (Analysis) & Var (Observed) & $\hat{L}_{est} $ \\
\hline
2  & 1.3423 & 1.3876 & 1.3452 (0.0031) & 2.047 (0.0036)\\
3  & 0.8221 & 0.8326 & 0.8191 (4.2e-5) & 3.041 (0.0001)\\
4  & 0.5907 & 0.5947 & 0.5941 (0.0059) & 4.004 (0.0351)\\
5  & 0.4607 & 0.4625 & 0.4622 (0.0010) & 5.003 (0.0099)\\
6  & 0.3774 & 0.3784 & 0.3800 (0.0009) & 5.977 (0.0131)\\
7  & 0.3196 & 0.3202 & 0.3188 (0.0037) & 7.029 (0.0769)\\
8  & 0.2771 & 0.2775 & 0.2786 (0.0002) & 7.970 (0.0065)\\
9  & 0.2446 & 0.2449 & 0.2455 (0.0007) & 8.979 (0.0244)\\
10 & 0.2189 & 0.2191 & 0.2183 (0.0012) & 10.035 (0.0538)\\
11 & 0.1981 & 0.1982 & 0.1975 (0.0001) & 11.038 (0.0063)\\
12 & 0.1809 & 0.1809 & 0.1810 (0.0022) & 12.001 (0.1419)\\
13 & 0.1664 & 0.1665 & 0.1661 (0.0002) & 13.031 (0.0191)\\
14 & 0.1541 & 0.1542 & 0.1534 (0.0016) & 14.068 (0.1387)\\
15 & 0.1435 & 0.1435 & 0.1432 (0.0019) & 15.036 (0.1967)\\
16 & 0.1342 & 0.1343 & 0.1342 (0.0009) & 16.006 (0.0987)\\
17 & 0.1261 & 0.1261 & 0.1249 (0.0005) & 17.159 (0.0636)\\
18 & 0.1189 & 0.1189 & 0.1192 (0.0010) & 17.959 (0.1406)\\
19 & 0.1125 & 0.1125 & 0.1126 (0.0004) & 18.976 (0.0788)\\
20 & 0.1067 & 0.1067 & 0.1074 (0.0005) & 19.889 (0.0937)\\
21 & 0.1015 & 0.1015 & 0.1017 (0.0002) & 20.975 (0.0357)\\
22 & 0.0968 & 0.0968 & 0.0970 (0.0002) & 21.952 (0.0337)\\
23 & 0.0925 & 0.0925 & 0.0924 (0.0006) & 23.027 (0.1386)\\
24 & 0.0886 & 0.0886 & 0.0887 (0.0002) & 23.957 (0.0507)\\
25 & 0.0849 & 0.0849 & 0.0846 (0.0007) & 25.094 (0.1934)
\end{tabular}
\caption{ Speckle Suppression Power: ENL and Variance }
\label{tab:enl_in_log_domain}
\end{table}

The finding here may also help to explain other seemingly unrelated results. For example, 
	Solbo \cite{Solbo_2006_TGRS} used standard deviation in the log-transformed domain to measure homogeneity, 
	while Lopes \cite{Lopes_TGRS_1990} proposed the use of variation co-efficient index $C_v = std(I)/avg(I)$ 
	to evaluate scene heterogeneity.

\subsubsection{Using log-variance to evaluate speckle filters}

Fig. \ref{fig:log_consistency_model} plots the histograms of homogenous SAR data over different radiometric values.
Both single-look simulated and multi-look processed/box-car filtered data is displayed.
Specifically the plots show that the log-transformed domain is consistent while the plots from the original 
domain are not.

\begin{figure}[h]
\centering  
\begin{tabular}{c}
	\subfloat[Single Look (Intensity)]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/orig_inconsistency_none.png.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Multi Look (Intensity)]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/orig_inconsistency_boxcar.png.eps} 	
		 \label{intensity}
	} \\
	\subfloat[Single-Look in Log Domain]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/log_consistency_none.png.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Multi-Look in Log Domain]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/log_consistency_boxcar.png.eps} 	
		 \label{intensity}
	} 
\end{tabular}
%}
\caption{The Inconsistency evident in the original SAR domain and the Emerged Consistency demonstrated in the log-transformed domain}
\label{fig:log_consistency_model}
\end{figure}

Fig. \ref{fig:log_consistency_filters} shows that all of the standard filters (Lee, Kuan, Frost and Gamma MAP) 
preserve this consistency in their filtered output. Intuitively, as the boxcar filter is actually similar to 
multi-look processing, thus its output also does exhibit the this consistency.
This consistency will also lead to the consistent sense of distance described earlier, which is significant 
because it is the tell-tale indicator of the consistent contrast and variance.
This ensures applicability of various target detection/classification algorithms 
which employ statistical properties in the un-filtered data, such as the ratio based discriminator 
in the original domain or the differential based discriminator in the log-transformed domain.

\begin{figure}
\centering  
\begin{tabular}{c}
	\subfloat[Lee filter]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/log_consistency_lee.png.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Kuan Filter]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/log_consistency_kuan.png.eps} 	
		 \label{intensity}
	} \\
	\subfloat[Frost Filter]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/log_consistency_frost.png.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Gamma MAP filter]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/log_consistency_map.png.eps} 	
		 \label{intensity}
	} 
\end{tabular}
%}
\caption{Filtered results: consistency in log-transformed domain}
\label{fig:log_consistency_filters}
\end{figure}

\begin{table}
\centering
\begin{tabular}{c|c|r|c|r}
Filter & Set & Log-Variance & $\hat{L}_{est}$ & Avg Intensity \\
\hline
none & 1 & 3.4149 (0.0003) & 1.1095 (5.8e-5) & 3.4795 (0.0029)\\
pde & 1 & 1.2674 (0.0014) & 2.1423 (0.0018) & 3.4806 (0.0029)\\
map & 1 & 1.2651 (0.0055) & 2.1454 (0.0072) & 3.1627 (0.0002)\\
lee & 1 & 0.4604 (0.0027) & 5.0206 (0.0268) & 3.4835 (0.0022)\\
kuan & 1 & 0.2979 (0.0024) & 7.4864 (0.0574) & 3.4748 (0.0054)\\
frost & 1 & 0.2852 (0.0012) & 7.7975 (0.0299) & 3.4799 (0.0027)\\
boxcar & 1 & 0.2513 (0.0011) & 8.7819 (0.0353) & 3.4799 (0.0027)\\
\hline
none & 2 & 3.4191 (0.0068) & 1.1088 (0.0012) & 9.4925 (0.0049)\\
pde & 2 & 1.6071 (0.0026) & 1.7952 (0.0021) & 9.4930 (0.0048)\\
map & 2 & 1.2646 (0.0023) & 2.1459 (0.0029) & 8.6091 (0.0162)\\
lee & 2 & 0.4653 (0.0026) & 4.9731 (0.0254) & 9.4936 (0.0080)\\
kuan & 2 & 0.2999 (0.0026) & 7.4391 (0.0602) & 9.4761 (0.0014)\\
frost & 2 & 0.2867 (0.0005) & 7.7595 (0.0128) & 9.4912 (0.0039)\\
boxcar & 2 & 0.2532 (0.0006) & 8.7203 (0.0193) & 9.4914 (0.0040)\\
\hline
none & 3 & 3.4203 (0.0187) & 1.1068 (0.0033) & 25.6234 (0.0389)\\
pde & 3 & 2.0353 (0.0021) & 1.5226 (0.0010) & 25.6236 (0.0389)\\
map & 3 & 1.2799 (0.0091) & 2.1263 (0.0116) & 23.2628 (0.0304)\\
lee & 3 & 0.4633 (2.4e-5) & 4.9926 (0.0002) & 25.6406 (0.0488)\\
kuan & 3 & 0.2964 (0.0010) & 7.5225 (0.0244) & 25.6112 (0.0226)\\
frost & 3 & 0.2862 (0.0008) & 7.7714 (0.0202) & 25.6263 (0.0409)\\
boxcar & 3 & 0.2529 (0.0005) & 8.7296 (0.0178) & 25.6261 (0.0407)
\end{tabular}
\caption{ Filters' Performance: Homogeneous Area }
\label{tab:homogeneous_performance_filters}
\end{table}

Table \ref{tab:homogeneous_performance_filters} provides a quantitative comparison among the filters.
It shows that the filters can all preserve the underlying radiometric values, while their speckle suppression power can be equivalently measured using either variance in the log-transformed domain, or the standard ENL index.

\subsubsection{ Estimating ENL from MSE index in homogeneous areas }

Over homogeneous area, the ground-truth is unchanged, i.e. $X^L_i=X^L \forall i$.
Assuming the filters are unbiased, or the bias is already compensated, i.e. $E(Y^L)=X^L$, then the MSE becomes $\Psi = \frac{ \sum^N_{i=1}{ (Y^L_i - E(Y^L)) } }{N} = var(Y^L)$.
That is, for known homogeneous scenes, MSE can be estimated as the observable variance of the filtered output in log-transformed domain.

Let us consider the speckle suppression effect of multi-look processing in the log-transformed domain. 
Since multi-look processing is unbiased, we will take the variance of log-transformed multi-look processing output as the MSE performance index. 
Hoek \cite{Hoekman_1991_TGRS} and Xie \cite{Xie_2002_TGRS} have given the variance for L-look log-transformed random variable as: $var(Y^L)= \frac{1}{\ln^2(2)} \left( \frac{\pi^2}{6} - \sum^{L-1}_{i=1}{\frac{1}{i^2}} \right)$.

Next we will be showing that the ENL, i.e. L, can be estimated from a given $var(Y^L)$. 
Taking results of the Euler proof for the Basel problem we have $\frac{\pi^2}{6} = \sum^{\infty}_{i=1}{ \frac{1}{i^2} } $, then $\Psi= \frac{1}{\ln^2(2)} \left( \sum^{\infty}_{i=L}{ \frac{1}{i^2} } \right) $.
Noting that $ \frac{1}{i} - \frac{1}{i+1} = \frac{1}{i(i+1)} < \frac{1}{i^2} < \frac{1}{i(i-1)} = \frac{1}{i-1} - \frac{1}{i}$, then $ \frac{1}{L} - \frac{1}{\infty} < \sum^{\infty}_{i=L}{ \frac{1}{i^2} }  < \frac{1}{L-1} - \frac{1}{\infty} $.
In fact, we could estimate:
\begin{equation}
  \Psi = \frac{1}{(L-0.5) \ln^2(2) }
\label{eqn:perf_index_analytic}
\end{equation}
Thus, the Equivalent Number of Look, i.e. L, can be estimated as:
\begin{equation}
ENL = \frac{1}{\Psi \ln^2(2)} + 0.5
\label{eqn:enl_analytic}
\end{equation}

From the section above, it can be shown that logarithmic transformation of single-look intensity, i.e. $X^L_1$, has a speckle power MSE of $\Psi_1 = var(X^L_1)=\frac{1}{\ln2^2} \frac{\pi^2}{6}$. 
With each additional 1-look in SAR multi-look processing, the speckle power will reduce by:  
\begin{equation}
 \Psi_n = \Psi_{n-1} - \left( \frac{1}{\ln2(n-1)} \right)^2 
\end{equation}
with $n$ being the number of look being averaged.

To verify the above analysis an experiment is carried out.
In the experiment, a known $256\times256$ homogeneous area is generated and is corrupted with SAR-speckle PDF. 
Different multi-look processing filters are repeatedly applied to the patch. 
A performance index is calculated for each simulation with their mean and variances reported. 
The analysis values are calculated from Equation \ref{eqn:perf_index_analytic}, together with estimated ENL from Equation \ref{eqn:enl_analytic}

\begin{table}
\centering

\begin{tabular}{c|c|c|c|c}
No  & Simulation                      &Analysis   &Est ENL       & Perf Index Improved \\% &85\%             &90\%\\
\hline
2   &1.3397 ($0.1036 \cdot 10^{-03}$) &1.3423	 &2.0536	& - \\
3   &0.8201 ($0.0188 \cdot 10^{-03}$) &0.8220	 &3.0379	& 0.5196\\
4   &0.5898 ($0.0217 \cdot 10^{-03}$) &0.5907	 &4.0287	& 0.2303\\
5   &0.4595 ($0.0067 \cdot 10^{-03}$) &0.4607	 &5.0292	& 0.1303\\
6   &0.3764 ($0.0013 \cdot 10^{-03}$) &0.3774	 &6.0291	& 0.0831\\
7   &0.3195 ($0.0073 \cdot 10^{-03}$) &0.3196	 &7.0149	& 0.0569\\
8   &0.2773 ($0.0043 \cdot 10^{-03}$) &0.2771	 &8.0056	& 0.0422\\
9   &0.2448 ($0.0016 \cdot 10^{-03}$) &0.2446	 &9.0010	& 0.0325\\
10  &0.2193 ($0.0015 \cdot 10^{-03}$) &0.2189	 &9.9908	& 0.0255\\
11  &0.1975 ($0.0024 \cdot 10^{-03}$) &0.1981	 &11.0400	& 0.0218\\
12  &0.1807 ($0.0013 \cdot 10^{-03}$) &0.1809	 &12.0157	& 0.0168\\
13  &0.1663 ($0.0004 \cdot 10^{-03}$) &0.1664	 &13.0121	& 0.0144\\
14  &0.1544 ($0.0004 \cdot 10^{-03}$) &0.1541	 &13.9783	& 0.0119\\
15  &0.1435 ($0.0003 \cdot 10^{-03}$) &0.1435	 &15.0075	& 0.0109\\
16  &0.1342 ($0.0003 \cdot 10^{-03}$) &0.1342	 &16.0071	& 0.0093\\
17  &0.1263 ($0.0003 \cdot 10^{-03}$) &0.1261	 &16.9765	& 0.0079\\
18  &0.1189 ($0.0004 \cdot 10^{-03}$) &0.1189	 &18.0085	& 0.0074\\
19  &0.1126 ($0.0006 \cdot 10^{-03}$) &0.1125	 &18.9777	& 0.0063\\
20  &0.1070 ($0.0004 \cdot 10^{-03}$) &0.1067	 &19.9528	& 0.0056\\
21  &0.1016 ($0.0003 \cdot 10^{-03}$) &0.1015	 &20.9847	& 0.0054\\
22  &0.0969 ($0.0004 \cdot 10^{-03}$) &0.0968	 &21.9806	& 0.0047\\
23  &0.0924 ($0.0005 \cdot 10^{-03}$) &0.0925	 &23.0188	& 0.0045\\
24  &0.0885 ($0.0002 \cdot 10^{-03}$) &0.0886	 &24.0214	& 0.0039\\
25  &0.0850 ($0.0001 \cdot 10^{-03}$) &0.0849	 &24.9942	& 0.0035
\end{tabular}

\caption{ ENL: Boxcar Filter's Speckle Suppression Power }
\label{tab:enl_in_log_domain}
\end{table}

%To evaluate the effects of number of looks on performance index, the performance index values for each ENL is computed and listed in table \ref{tab:enl_in_log_domain}. 
%To compare our filter with boxcar filters, the error is reported not only as a single performance index value but also as the whole distribution, depicted as the histogram of error in log-domain, see figure ???. Homoskedastic effects should also make it possible to evaluate performance speckle filter in heterogenous scenes, however such evaluation is outside the scope of current paper.

\subsection{ Evaluating SAR Speckle Filters over Heterogeneous Scenes }
\label{sec:eval_hetero}

Similar to the drop in variance allowing one to estimate the effect of each extra look averages being used over homogeneous areas, rises in the error variance can also be used to gauge the effects of underlying spatial variation on speckle filtering effectiveness. 
In original SAR domain, evaluating speckle filters' response to different heterogeneous areas are difficult, due to heteroskedastic conditions.
In log-transformed domain, thanks to homoskedastic property, MSE will be shown to be relevant in evaluating speckle response to heterogeneous scenes (To be continued).
%An simple 4x4 edge is added into the 256x256 area above, filters are applied and MSE reported.

%\section{Evaluating Speckle Filters on Heterogenous Area}

%For the purpose of comparative evaluation among speckle filters, the evaluation methodology needs to be quantitative and repeatable.
%First, several test patterns is created to simulate and determine the performance of various speckle filters.
%Then the methodology to evaluate speckle filter in heterogeneous area using the MSE criteria in log-transformed domain is described.
%As the ground truth is available, the MSE criteria is validated by analysis and experimental results which show that lower MSE is correlated with better feature preservation.

In this section, the MSE criteria ($MSE_{true}$) is validated through experiments and analysis. 
From a statistical estimation framework point of view, the use of MSE to evaluate statistical estimators is natural.
Because the log-transformation converts heteroskedastic SAR speckle into a homoskedastic distribution model, 
	the log-transformed domains MSE is much preferred to MSE in the original SAR domain.
Experimentally, on clear-cut simple target-background patterns, the results that follow will indicate that
	the lower the MSE achievable by a speckle filter,
	the better the feature preservation performance recorded for its output.

%We believe that the qualitative requirement of speckle suppression can be quantified as the variance in 
%log-transformed domain. 
The general requirements of feature preservation, for simple scenes with only targets and clutter, can be broken 
down into the requirements of radiometric preservation and speckle suppression.
In the log-transformed domain, these smaller requirements are equivalent to the bias and variance evaluation of 
statistical estimators.
Overall, while the MSE index combines the measurements of bias and variance evaluation, 
	the feature preservation requirement, in the context of simple target and clutter scenes, can be measured by 
	the standard metric for target detectability: the Area Under the ROC Curve (AUC).
We show experimentally that MSE inversely correlates with the AUC index, for each of the simulated patterns.

Since the filters are expected to be consistently behaved in the log-transformed domain, as we repeat a pattern 
multiple times, the histograms of the target and background areas can be reliably obtained.
%They are expected to be consistent in log-transformed domain.
The target and background can then be seperated using a simple threshold based classification model.
The separability of the two stochastic populations are judged by the standard ROC, 
	and the quantitative and normalized metric of AUC can then be used as an evaluation metric.

The types of test pattern used in this paper include the followings:
  point targets, line targets, edge targets and a heterogeneous checker board. 
These test patterns could of course be concatenated into a larger composite test image, 
although we will consider them as separate images to allow easier analysis of the results.
%However we would prefer to highlight the results separately for each individual area.
Each pattern comprises two classes of ground-truth: background and target areas. We follow the convention used in the 
radar community where the target is signified by the brighter area of the image. 
Fig \ref{fig:hetero_patterns} shows a small section ($32 \times 32$ window) of each pattern.

\begin{figure}
\begin{tabular}{c}
	\subfloat[Line: each line is 2 pixels wide, separated by 6 pixels background]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/pattern_line2.png.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Edge: each stripe is 4 pixels in width]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/pattern_edge.png.eps} 	
		 \label{intensity}
	} \\
	\subfloat[Point: each point is a $2 \times 2$ square spacing 6 pixels apart]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/pattern_point.png.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Checker board: the squares are 4 pixels wide each side]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/pattern_checker.png.eps} 	
		 \label{intensity}
	} 
\end{tabular}
%}
\centering
\caption{Example windows of ground truth patterns, each $32 \times 32$ pixels in size.}
\label{fig:hetero_patterns}
\end{figure}

Large patches of these patterns ($512 \times 512$) are first corrupted with single look speckle. The 
filters are then applied onto these noised images.
From a high-level perspective, Fig. \ref{fig:hetero_patterns_roc_auc} shows that feature preservation can be 
evaluated by examining the separability of the two background and target populations.
Subfigures \ref{fig:hetero_patterns_roc_auc:hist_unfiltered} and \ref{fig:hetero_patterns_roc_auc:hist_kuan_filtered} 
	allows visual evaluation of target and clutter histograms and their separability.
Quantitative evaluation of these separability is carried out by plotting the Receiver Operating Curve (ROC), 
in subfigures \ref{fig:hetero_patterns_roc_auc:roc_unfiltered} and \ref{fig:hetero_patterns_roc_auc:roc_kuan_filtered} 
respectively, and is measured by computing the Area Under these Curves (AUC).

%visualizes the two histograms
%The separability of the two histograms, subfigures \ref{fig:hetero_patterns_roc_auc:hist_unfiltered} and \ref{fig:hetero_patterns_roc_auc:hist_kuan_filtered}, 
%	is visualized by plotting the Receiver Operating Curve (ROC), subfigures \ref{fig:hetero_patterns_roc_auc:roc_unfiltered} and \ref{fig:hetero_patterns_roc_auc:roc_kuan_filtered} respectively,
%	and is measured by computing the Area Under these Curves (AUC).

\clearpage

\begin{figure}
\centering  
\begin{tabular}{c}
	\subfloat[Simulated Image]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.edge.none.fi.jpg.eps} 	
		 \label{fig:hetero_patterns_roc_auc:amplitude}
	} 
	\hfill	
	\subfloat[Kuan Filtered Image]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.edge.kuan.fi.jpg.eps} 	
		 \label{fig:hetero_patterns_roc_auc:intensity}
	} \\
	\subfloat[Histograms: Unfiltered]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.histograms.edge.none.fi.png.eps} 	
		 \label{fig:hetero_patterns_roc_auc:hist_unfiltered}
	} 
	\hfill	
	\subfloat[Histograms: Kuan Filtered]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.histograms.edge.kuan.fi.png.eps} 	
		 \label{fig:hetero_patterns_roc_auc:hist_kuan_filtered}
	}  \\
	\subfloat[ROC: Unfiltered, AUC=0.738]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.roc_auc.edge.none.fi.png.eps} 	
		 \label{fig:hetero_patterns_roc_auc:roc_unfiltered}
	} 
	\hfill	
	\subfloat[ROC: Kuan Filtered, AUC=0.885]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.roc_auc.edge.kuan.fi.png.eps} 	
		 \label{fig:hetero_patterns_roc_auc:roc_kuan_filtered}
	} 
\end{tabular}
\caption{Target and Clutter Separability: Histograms and resulting ROC Curve visualisation}
\label{fig:hetero_patterns_roc_auc}
\end{figure}

Detailed examination of figures \ref{fig:hetero_patterns_roc_auc} and \ref{fig:hetero_patterns_mse} 
	can help to explain how the MSE is related to the histograms' separation capability.
In pre-filtered images, shown in subfigures \ref{fig:hetero_patterns_roc_auc:hist_unfiltered} and 
\ref{fig:hetero_patterns_mse:unfiltered}, there is no bias error visible.
Then the seperability of clutter and target populations depends only on the variance of the additive noise. 
This variance is visualized as the horizontal spread of the histograms.
Naturally, given a fixed location (i.e. expectation) of the two populations, the smaller the spread (i.e. variance), 
the better the separation capability.

In post-filtered images, shown in subfigures \ref{fig:hetero_patterns_roc_auc:hist_kuan_filtered} and 
\ref{fig:hetero_patterns_mse:kuan_filtered}, the situation is more complicated.
Here, besides the effect of the histogram spread, one also needs to take into account the bias error.
Close investigation of subfigure \ref{fig:hetero_patterns_mse:kuan_filtered} indicates that the output of the 
Kuan filter, and in fact the outputs of all other filters (which are not reproduced here due to space constraints), 
also introduce bias errors.
Specifically, the target (brighter) populations are always under-estimated and the clutter (darker) population are 
always over-estimated. This is probably due to the entropy reduction effect of the speckle filters.
Apparently, assuming that the variances are fixed, the lower these bias errors, the better the separation capability.

%Also the speckle suppression is related to the measured variance.
Thus, the MSE performance of the estimator, which combines the effect of bias and variance error, can be used to 
evaluate the separability of the two histograms. 
Table \ref{tab:mse_auc_in_log_domain} provides the measurements of MSE and AUC performance for various filters and 
patterns.
It shows that the MSE is inversely correlated to this separability index.
The first column in Table \ref{tab:mse_auc_corr_coeff} quantitatively measures this statistical correlation.
The results suggest that the lower MSE achievable by the filters would, in general, lead to better feature preservation.

\begin{figure}
\centering  
\begin{tabular}{c}
	\subfloat[Error Image: Unfiltered]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.edge.none.gt.jpg.eps} 	
		 \label{fig:hetero_patterns_mse:amplitude}
	} 
	\hfill	
	\subfloat[Error Image: Kuan Filtered]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.edge.kuan.gt.jpg.eps} 	
		 \label{fig:hetero_patterns_mse:intensity}
	} \\
	\subfloat[Error Histograms: Unfiltered]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.histograms.edge.none.gt.png.eps} 	
		 \label{fig:hetero_patterns_mse:unfiltered}
	} 
	\hfill	
	\subfloat[Error Histograms: Kuan-filtered]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/heterogenous_patterns.histograms.edge.kuan.gt.png.eps} 	
		 \label{fig:hetero_patterns_mse:kuan_filtered}
	}  
\end{tabular}
\caption{Bias Error Investigation: Image and Histogram Visualisation}
\label{fig:hetero_patterns_mse}
\end{figure}

\begin{table}
\centering
\begin{tabular}{c|r|c|c|r}
Pattern  & Filter  & AUC & $MSE_{true}$   & $MSE_{noise}$      \\% &85\%             &90\%\\
\hline
 point  &  unfiltered  &  0.741 (0.7e-3)  &  4.101 (1.3e-3)  &  2e-33 (4e-36) \\
 point  &  pde  &  0.789 (1.3e-3)  &  1.291 (5.1e-3)  &  1.817 (2.2e-3)\\
 point  &  map  &  0.813 (1.4e-3)  &  1.679 (2.2e-3)  &  2.183 (7.4e-3)\\
 point  &  frost  &  0.836 (1.8e-3)  &  0.536 (2.4e-3)  &  4.976 (6.9e-3)\\
 point  &  lee  &  0.857 (0.9e-3)  &  0.615 (3.1e-3)  &  3.189 (0.9e-3)\\
 point  &  boxcar  &  0.871 (1.5e-3)  &  0.471 (1.7e-3)  &  4.503 (5.4e-3)\\
 point  &  kuan  &  0.882 (1.1e-3)  &  0.448 (2.3e-3)  &  3.859 (3.4e-3)\\
\hline
 edge  &  unfiltered  &  0.738 (1.5e-4)  &  4.128 (9.0e-3)  &  2e-33 (1e-35)\\
 edge  &  pde  &  0.783 (0.6e-4)  &  1.409 (0.3e-3)  &  1.589 (0.4e-2)\\
 edge  &  map  &  0.830 (0.6e-4)  &  1.712 (3.5e-3)  &  2.184 (0.8e-2)\\
 edge  &  frost  &  0.841 (0.6e-4)  &  0.551 (0.1e-3)  &  5.035 (1.3e-2)\\
 edge  &  boxcar  &  0.871 (0.2e-4)  &  0.486 (0.2e-3)  &  4.560 (1.3e-2)\\
 edge  &  lee  &  0.872 (0.7e-4)  &  0.619 (2.2e-3)  &  3.233 (1.3e-2)\\
 edge  &  kuan  &  0.885 (0.4e-4)  &  0.471 (2.0e-3)  &  3.909 (1.1e-2)\\
\hline
 checker  &  unfiltered  &  0.738 (4.5e-4)  &  4.11 (7.8e-3)  &  2e-33 (5e-36)\\
 checker  &  pde  &  0.785 (6.2e-4)  &  1.445 (2.6e-3)  &  1.586 (2.4e-3)\\
 checker  &  map  &  0.836 (4.0e-4)  &  1.663 (1.1e-3)  &  2.213 (4.6e-3)\\
 checker  &  frost  &  0.855 (5.6e-4)  &  0.528 (2.2e-3)  &  4.965 (9.3e-3)\\
 checker  &  lee  &  0.879 (2.2e-4)  &  0.605 (1.7e-3)  &  3.229 (2.8e-3)\\
 checker  &  boxcar  &  0.883 (6.4e-4)  &  0.466 (2.4e-3)  &  4.493 (9.1e-3)\\
 checker  &  kuan  &  0.894 (4.2e-4)  &  0.453 (0.9e-3)  &  3.860 (9.5e-3)\\
\hline
 line  &  unfiltered  &  0.737 (1.1e-3)  &  4.129 (7.3e-3)  &  2e-33 (6e-36)\\
 line  &  pde  &  0.752 (1.2e-3)  &  1.339 (2.0e-3)  &  1.885 (3.7e-3)\\
 line  &  map  &  0.801 (1.6e-3)  &  1.706 (5.8e-3)  &  2.188 (3.5e-3)\\
 line  &  frost  &  0.831 (1.3e-3)  &  0.551 (1.1e-3)  &  5.023 (0.5e-3)\\
 line  &  lee  &  0.847 (1.5e-3)  &  0.623 (2.9e-3)  &  3.228 (4.8e-3)\\
 line  &  boxcar  &  0.865 (1.3e-3)  &  0.486 (0.9e-3)  &  4.549 (0.9e-3)\\
 line  &  kuan  &  0.874 (1.9e-3)  &  0.464 (1.9e-3)  &  3.897 (4.9e-3)\\
\hline
\end{tabular}

\caption{Lower MSE suggest better feature detection, measured by the AUC index}
\label{tab:mse_auc_in_log_domain}
\end{table}

\begin{table}
\centering
\begin{tabular}{c|c|c}
Pattern  & AUC - $MSE_{true}$  & AUC - $MSE_{benchmark}$  \\
\hline
edge & -0.8958 (1.5e-05) &   -0.9778  (1.6e-09) \\
point &     -0.9012   (1.1e-05)   &    -0.9816        (5.3e-10) \\
checker &   -0.9077     (7.3e-06)  &  -0.9829       (3.5e-10) \\
line &      -0.8223     (3.1e-04)  &   -0.9421       (4.8e-07) \\
\hline
\end{tabular}

\caption{The correlation between MSE and AUC evaluation criteria (inside the brackets are corresponding p-values)}
\label{tab:mse_auc_corr_coeff}
\end{table}

\subsection{Using MSE to find the most suitable speckle filter for practical SAR images}
\label{sec:practical_conjecture}

In this section, our conjecture
%(??? you want to emphasize this is a proposal that can't be proven mathematically ?)
%ans: it is true that the conclusion cannot be proved, due to lack of absolute ground truth in real SAR images
%our conclusion however is 1. extended from a simulations which is believed to be in agreement with realities.
of using MSE in the log-transformed domain to find the most suitable speckle filter 
for practical real-captured SAR images is described.
In these scenarios, the ground-truth, and hence the true MSE, is not available.
Therefore, only the residential MSE is computable.
Assuming the level of speckle noise (i.e. ENL or $MSE_{base}$) is known or can be estimated, then the benchmark 
MSE is also measurable.	
Our heuristic rule is that the best filtered results are those that have minimal benchmarked MSE. 
This heuristic rule allow us to choose the ``best'' filtered results from an array of standard speckle filters 
for a captured SAR image, where the ground-truth and hence true MSE is not available.
Intuitively, the hypothesis is that the best speckle filter for a given scene is the one having its 
removed variation being closest to the inherent speckle noise.

Experimental results are presented as empirical evidence supporting the conjecture.
The experiments in the previous sections are repeated on single-look SAR images 
	which are simulated from given ground-truth aerial images. 
%The a heuristic rule is that 
%	if the best filtered results are those with minimal MSE 
%	then just by observing the residual MSE, 
%	the observable measure of these results will also achieve 
%		being the closest to the MSE of the inherent noise.
%Besides using visual evaluation to validate the conjecture, 
%	the full justification of this conjecture is outside the scope of this paper.
%Real images however are more complex than the patterns illustrated above. 
Then the filters are applied onto the simulated SAR images. 
Fig. \ref{fig:real_simulated_images} illustrates some of the images used for our experiments.

\begin{figure}
\centering  
\begin{tabular}{c}
	\subfloat[A Rural Area in Vietnam]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/simulated_images.vietnam_rural.gt.jpg.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[A Suburb of Ha Noi]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/simulated_images.hanoi_suburb.gt.jpg.eps} 	
		 \label{intensity}
	} %\\
%	\subfloat[NTU Campus]{
%		 \epsfxsize=1.5in
%		 \epsfysize=1.5in
%		 \epsffile{src/Aerialcampus.eps} 	
%		 \label{amplitude}
%	} 
%	\hfill	
%	\subfloat[Chu Thap Island of Vietnam]{
%		 \epsfxsize=1.5in
%		 \epsfysize=1.5in
%		 \epsffile{src/fiery.eps} 	
%		 \label{intensity}
%	} 
\end{tabular}
\caption{Ground Truth Images for simulation}
\label{fig:real_simulated_images}
\end{figure}

With the use of MSE being validated from previous experiments, 
	the most suitable filter can be considered as the one with the lowest true MSE.
Table \ref{tab:mse_true_noise_log_domain} shows that among the various filters used, 
	the most suitable filter is also the filter that 
		has its observable residual MSE value being closest to the noise MSE (4.1167 in the case of this example).
%We validate the idea by qualititative evaluation, an example of 
The conjecture is also validated using visual evaluation, an example of
	which is presented in Fig. \ref{fig:real_simulated_image_results}.

\begin{table}
\centering
\begin{tabular}{r|r|c|c}
Pattern  & Filter  & $MSE_{true}$   & $MSE_{noise}$      \\% &85\%             &90\%\\
%\hline
%chu thap island	& none		& 4.1066	& 1.2e-33\\
%chu thap island	& pde			& 1.7966	& 1.0866\\
%chu thap island	& lee			& 0.5772	& 3.2378\\
%chu thap island	& frost		& 0.4701	& 4.8608\\
%chu thap island	& kuan		& 0.4196	& 3.7945\\
%chu thap island	& boxcar	& 0.4165	& 4.3952\\
%\hline
%ntu campus	& none		& 4.1231	& 8.7e-35\\
%ntu campus	& pde			& 3.6598	& 0.0733\\
%ntu campus	& lee			& 0.6335	& 3.2673\\
%ntu campus	& frost		& 0.5776	& 5.0569\\
%ntu campus	& boxcar	& 0.5136	& 4.5839\\
%ntu campus	& kuan		& 0.4910	& 3.9444\\
\hline
Vietnam rural	& none		& 4.1174	& 4e-35\\
Vietnam rural	& pde			& 3.8022	& 0.0368\\
Vietnam rural	& lee			& 0.4984	& 3.2555\\
Vietnam rural	& frost		& 0.3490	& 4.6856\\
Vietnam rural	& kuan		& 0.3396	& 3.6877\\
Vietnam rural	& boxcar	& 0.3107	& 4.2328\\
\hline
Hanoi suburb	& none		& 4.1321	& 4e-35\\
Hanoi suburb	& pde			& 3.8004	& 0.0391\\
Hanoi suburb	& lee			& 0.5261	& 3.2598\\
Hanoi suburb	& frost		& 0.3811	& 4.7427\\
Hanoi suburb	& kuan		& 0.3619	& 3.7270\\
Hanoi suburb	& boxcar	& 0.3395	& 4.2882\\
\hline
\end{tabular}

\caption{If the best filters are the ones with smallest true MSE, then their observable noise-MSE are also the ones closest to the MSE of inherent noise}
\label{tab:mse_true_noise_log_domain}
\end{table}

%\begin{figure*}
\begin{figure}
\normalsize
\begin{center}
\begin{tabular}{c}
	\subfloat[Ground-Truth Image ]{
		 \epsfxsize=1.6in
		 \epsfysize=1.6in
		 \epsffile{images/simulated_images.vietnam_rural.gt.jpg.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Unfiltered Image $MSE_{true}=MSE_{base}=4.1174$]{
		 \epsfxsize=1.6in
		 \epsfysize=1.6in
		 \epsffile{images/simulated_images.vietnam_rural.none.fi.jpg.eps} 	
		 \label{intensity}
	} \\
	\subfloat[PDE Result: $MSE_{true}=3.8022,MSE_{noise}=0.0073$]{
		 \epsfxsize=1.6in
		 \epsfysize=1.6in
		 \epsffile{images/simulated_images.vietnam_rural.pde.fi.jpg.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Lee Result: $MSE_{true}=0.4984,MSE_{noise}=3.25553$]{
		 \epsfxsize=1.6in
		 \epsfysize=1.6in
		 \epsffile{images/simulated_images.vietnam_rural.lee.fi.jpg.eps} 	
		 \label{intensity}
	} \\
	\subfloat[Frost Result: $MSE_{true}=0.3490, MSE_{noise} = 4.6856$]{
		 \epsfxsize=1.6in
		 \epsfysize=1.6in
		 \epsffile{images/simulated_images.vietnam_rural.frost.fi.jpg.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Boxcar Result: $MSE_{true} = 0.3107, MSE_{noise}= 4.2328$]{
		 \epsfxsize=1.6in
		 \epsfysize=1.6in
		 \epsffile{images/simulated_images.vietnam_rural.boxcar.fi.jpg.eps} 	
		 \label{intensity}
	}
\end{tabular}

\caption{Filtering Simulated Real Images: Qualititative Validation}
\label{fig:real_simulated_image_results}
\end{center}
\end{figure}
%\end{figure*}

The results from the experiments in previous sections can also be used to validate the use of residual and 
benchmark MSE.
The AUC-$MSE_{benchmark}$ column in Table \ref{tab:mse_auc_corr_coeff} shows that the criteria index is strongly 
correlated with feature classification capability, measured by the standard AUC index.

The conjecture is also validated in real SAR images.
Different speckle filters are applied onto a real RADARSAT-2 image.
In this case, since the ground-truth is not available, 
	only visual evaluation can be used to validate our conjecture.
Table \ref{tab:mse_in_real_image} tabulates the computed MSE of the ``removed'' additive noise.
Evidently all filters still leave some noise ``unremoved'', in which case, the higher removed noise MSE 
would probably suggest a more suitable filter.
Fig. \ref{fig:real_image_results} tends to confirm this visually.
%IVMThe conclusion in this can be found through further experiments with more images, 
%IVM	which cannot be shown here due to space limitation.

\begin{table}
\centering
\begin{tabular}{c|r|r}
Filter & $MSE_{residual}$ & $MSE_{benchmark}$\\
\hline
pde & 0.2583 & 3.8584 \\
map & 2.6936 & 1.4231 \\
kuan & 3.3924 & 0.7243 \\
lee & 3.4172 & 0.6995 \\
boxcar & 3.6918 & 0.4249 \\
frost & 4.1191 & 0.0024 \\
\hline
none & $MSE_{base}$ & 4.1167 
\end{tabular}
\caption{Our Conjecture: Most Suitable Speckle Filter For The Scene Can Be Chosen Using The Residual MSE.}
\label{tab:mse_in_real_image}
\end{table}

\begin{figure}
\begin{tabular}{c}
	\subfloat[PDE Filter: $MSE_{benchmark}=3.8584$]{
		 \epsfxsize=3cm
		 \epsfysize=3cm
		 \epsffile{images/heterogenous_real.log.image.pde.jpg.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[MAP Filter: $MSE_{benchmark}=1.4231$]{
		 \epsfxsize=3cm
		 \epsfysize=3cm
		 \epsffile{images/heterogenous_real.log.image.map.jpg.eps} 	
		 \label{intensity}
	} \\
	\subfloat[Lee Filter: $MSE_{benchmark}=0.6995$]{
		 \epsfxsize=3cm
		 \epsfysize=3cm
		 \epsffile{images/heterogenous_real.log.image.lee.jpg.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[Frost Filter: $MSE_{benchmark}=0.0024$]{
		 \epsfxsize=3cm
		 \epsfysize=3cm
		 \epsffile{images/heterogenous_real.log.image.frost.jpg.eps} 	
		 \label{intensity}
	}
\end{tabular}
\caption{Filtering Real Images: Smaller $MSE_{benchmark}$ suggests visually better images}
\label{fig:real_image_results}
\end{figure}

\subsection{Discussion and Conclusions}

\subsubsection{Discussion}

Although it is widely known that log transformation transforms multiplicative SAR speckle into additive noise, 
	one should note that the noise is not Gaussian. In fact, figures presented in the previous sections show that 
	they are not even centered around the origin. 
This may explain why averaging filters in the log-transformed domain (e.g. \cite{Arsenault_JOptSocAm_1976}) do not 
	work very well in practice.
To counter this, the use of maximum likelihood estimation, instead of simple averaging, 
	is suggested \cite{Le_2011_ACRS}. Interestingly, averaging is also the MLE operator in the SAR's original domain.

Log transformation also brings about a few consistent measures of dissimilarity.
	This consistency can be found in single-look or multi-look SAR data, 
	as well as in filtered outputs of various ``standard'' speckle filters.
This allows a variety of target detection/classification algorithms
  - which exploit these consistent statistical properties -
  to not only perform brilliantly on pre-filtered data,
  but also be found practically working on post-filtered data.

These consistent measures of distance in the log-transformed domain could probably have implications beyond speckle filtering.
For example, in the subsequent tasks of designing target detectors or classifiers,
  it is normally desirable for the solution to work on both unfiltered and post-filtered SAR data.
In such cases, these consistent dissimilarity measures found in the log-transformed domain could provide a sound theoretical
	basis. In fact, a number of proposed solutions have already appeared to exploit this feature. 
A case in point is the simple ratio based discriminator in the original SAR domain.
Looking backwards, in designing new speckle filters,
  by ensuring that the filtered output preserves these consistent properties,
	the newly designed filters would be eligible to be employed as pre-processing step for properly designed classifiers / detector.
From a higher level perspective, the applicability of Gauss-Markov theorem and consequently a meaningful MSE 
  predict fruitful applications for a variety of least-squared-error algorithms in the homoskedastic log-transformed domain.

Of course, there are other speckle filters that do not preserve such consistency 
	(e.g. the PDE filter \cite{You_TIP_2000}). One could argue that it would not be fair to judge such filters 
	using the MSE criteria, which tend to favour the ``standard'' filters.
While we respect any other criteria that have been used, we reiterate the two salient points of our approach. 
Firstly, our MSE criteria is closely related to the basic ENL criteria.
As shown in previous sections, experimental results indicated that
  the ENL measures for such filter (i.e. the PDE filter) differ depending  on the radiometric values. 
Secondly, speckle filters need to serve a purpose, and evaluation criteria should be relevant to such purpose.
The MSE criteria is shown to be related to feature preservation requirements.
Thus it is relevant towards subsequent target detection / classification processing, 
	believed to be a common subsequent processing step. 
For these two reasons, the MSE criteria is advocated.

In the experiments above, speckle filters with a $3 \times 3$ sliding window were used 
	even though we are aware that the normal window size employed is much larger. 
The reasons for maintaining such a small window is that 
	smaller-sized filters facilitate the use of smaller patterns without too much concern for crosstalk 
	among adjacent targets.
In addition, we focused on the use of MSE in the log-transformed domain for the evaluation of speckle filters, 
	and do not wish to advocate any particular ``best'' filter. 
In other words, we addressed the methodology of evaluating speckle filters, 
	and did not directly address the design of speckle filters.
However, interested filter designers are invited to download the open Matlab source code used in this paper 
for evaluating their own designs
  \footnote{This can be found at \texttt{http://www.lintech.org/Hai/Matlab}}.

In this paper, stochastic simulation is used extensively to evaluate the performance of statistical estimators 
(i.e. speckle filters).
The use of small and simple patterns allows detailed analysis
  which can then be done repeatedly and reliably against the 
stochastic nature of SAR data.
This also helps in mapping qualitative requirements for the speckle removal process into specific and quantitative 
requirements in the design of speckle filters.
Its use also provides an absolute ground-truth as a solid base for comparison of results across different papers.

The main drawback, of course, is that ground truth often does not exist in real-captured SAR images.
Thus the result extension towards real images is only analogical.
While the proposed rule is heuristic, 
	the experimental results presented in this paper are shown empirically to be valid.

We distinctively divided speckle filtering into two distinct scenarios, that of homogeneity and heterogeneity.
While perfect homogeneous ground truth can be defined, 
	different heterogeneous patterns exhibits different levels of heterogeneity. 
As such, different measures have been proposed to evaluate heterogeneity levels.
Unfortunately, for real captured images, where the absolute ground-truth is not available, 
	there is no known and certain way to assert a given image as being perfectly homogeneous or 
	absolutely heterogeneous.
Thus while the distinction helps in clarifying the concepts, 
	it is probably a leaky abstraction.

The patterns used are chosen based on our experience, which may affects evaluation results. 
As different patterns result in different homogeneity/heterogeneity degrees, 
	they also appear to affect the performance and ranking of each speckle filters.
One extreme example is that of perfect homogeneity, where boxcar filters would exhibit a respectable performance.
While, at the other end of the spectrum: that of high heterogeneity, it is common knowledge that the boxcar filter 
	would not perform that well.

The current scheme of finding the most suitable speckle filter requires the application of all filters before 
a decision is made, which probably demands excessive processing. 
A possibly better alternative would be to predict the choice, escaping such massive requirements of computational power.
This, however, is outside the scope of this paper.

\subsubsection{Conclusion}

To summarise, speckle filters are generally evaluated using many different qualitative criteria.
To compare the filters against each other, a methodology is needed
  to quantify these qualitative requirements,
  and subsequently to measure, compare and evaluate these quantities in the filtered results.
Central to all these is the need for a consistent sense of distance.

Logarithmic transformation has been shown 
  not only to convert multiplicative and heteroskedastic noise in the original SAR domain to additive and homoskedastic values,
  but also to offer a few consistent measures of distance.
With the Gauss-Markov theorem becoming applicable in this domain, we describe and propose the use of MSE in the 
log-transformed domain as a unifying criteria to quantitatively measure different requirements for speckle filters.

Our contribution is mainly centered around a few points. 
Firstly, a mathematical equation is established to link the ENL index to the variance in the log-transformed domain. 
We also illustrate the use of log-variance evaluation -
  which is shown to be equivalent to the standard ENL evaluation -
  in evaluating the speckle suppression effect of different speckle filters.
Secondly we show that MSE is inversely correlated to the AUC index for simulated heterogeneous areas.
The result suggests that the smaller MSE a filter could achieve, the better it would
be at discriminating between the underlying radiometric features.
%The result suggests that the smaller MSE a filter could achieve, the easier it would be discriminating the underlying radiometric features.
Thirdly, the practical contribution is to suggest an heuristic rules using the benchmark MSE to find 
the most suitable speckle filter for any given scene. 
Combined, we propose the use of MSE in log-transformed domain in evaluating the performance of different speckle filters in a variety of evaluation scenarios, and suggest several evaluation methodologies that may be useful in this regard.

It should also be noted that similar consistent measures of distance also exist in polarimetric SAR (POLSAR) data. 
Thus future work may explore the applicability of MSE approaches to POLSAR data analysis and processing.

%\section{FMLE POLSAR Speckle Filter using the consistent measures}

%This section describes the application of one consistent measure of distance,
%  namely contrast, in establishing a new POLSAR speckle filter.
%The first sub-section illuminates how the principles of fuzzy logic and Maximum Liklyhood Estimation (MLE) can be employed.
%The second sub-section illustrates that FMLE estimation preserves the consistency of contrast,
%  and hence consistent variance and ENL can be expected.
%The third sub-section introduces the recursive FMLE POLSAR speckle filter.

%\subsection{Fuzzy MLE estimation of POLSAR covariance matrix}
%\subsection{Preservation of the Consistency Properties in the FMLE outputs}
%\subsection{Recursive FMLE Speckle Filters}
%Compare increasing size FMLE vs decreasing size FMLE

\section{Evaluating POLSAR Speckle Filters using the consistent measures}

This section describes how POLSAR filters can be evaluated
  using the consistent measures of distance found in the homoskedastic log-transformed domain.
The first sub-section theoretically illustrates how the consistent variance is linked to the ENL index.
The second sub-section then demonstrate how the POLSAR speckle filters can be pratically and quantitatively evaluated over homogeneous areas by measuring the observable consistent variance after the log-transformation.
The third sub-section finally illlustrates how POLSAR speckle filters can be visually evaluated on heterogeneous areas.

\subsection{POLSAR ENL Estimation using the Homoskedastic Property}

While Anfinsen solves the above equation to find ENL,
  we took variance approach.
Since we can ignore the constants $d\ln2 + d\ln{L} -ln|\Sigma|$ this approach might be simpler.

\begin{equation}
  var \left[ ln|C_v| \right] = \sum^{d-1}_{i=0} \psi^1(L-i)
\end{equation}

For SAR, d=1 and $C_v=I$ then
$var(\ln{I}) = \psi^1(L)=\sum^{\infty}_{i=L}1/i^2$.
Since $1/L < \sum^{\infty}_{i=L}1/i^2 < 1/(L-1)$ Then an approximation is made to compute L as
$var(\ln{I}) = 1/(L-0.5)$ or $L=1/var(\ln{I}) + 0.5 $

For partial POLSAR, d=2, then
\begin{equation}
  var \left[ ln|C_v| \right] = \psi^1(L) + \psi^1(L-1) \\
    = \sum^{\infty}_{i=L}1/i^2 + \sum^{\infty}_{i=L-1}1/i^2 \\    
\end{equation}
That is $1/L + 1/(L-1) < var_d \left[ \ln |C_v| \right] < 1/(L-1) + 1/(L-2)$.
Thus an approximation is made $var_d \left[ \ln |C_v| \right] = 2/(L-1)$
or $L = 2/var_d \left[ \ln |C_v| \right] + 1$

For full POLSAR, d=3, then
\begin{equation}
  var \left[ ln|C_v| \right] = \psi^1(L) + \psi^1(L-1) + \psi^1(L-2) \\
    = \sum^{\infty}_{i=L}1/i^2 + \sum^{\infty}_{i=L-1}1/i^2  + \sum^{\infty}_{i=L-2}1/i^2 \\    
\end{equation}
That is $1/L + 1/(L-1) + 1/(L-2) < var_d \left[ \ln |C_v| \right] < 1/(L-1) + 1/(L-2) + 1/(L-3)$.
With L>3, it is trivial to prove that $3/(L-1.5)$ also is bounded by the above thresholds.
Thus an approximation is made $var_d \left[ \ln |C_v| \right] = 3/(L-1.5)$
or $L = 3/var_d \left[ \ln |C_v| \right] + 1.5$

$L_{det}^2 = \ln{|Cv|}$
$var(L_{det}^2)= \frac{2}{ENL-1}$
$L_{est}=\frac{2}{var(L_{det}^2)} + 1$

\begin{table}[h!]
\centering
\begin{tabular}{r|c|c}
 ENL & log-var(std)    & L est (enl)      \\
\hline
   2 & 2.2962 (0.1415) & 1.8734 (0.0509)  \\
   3 & 1.0597 (0.0453) & 2.8899 (0.0812)  \\
   4 & 0.6543 (0.0369) & 4.0646 (0.1831)  \\
   5 & 0.4837 (0.0142) & 5.1373 (0.1239)  \\
   6 & 0.3945 (0.0159) & 6.0755 (0.2129)  \\
   7 & 0.3229 (0.0086) & 7.1966 (0.1640)  \\
   8 & 0.2942 (0.0187) & 7.8209 (0.4616)  \\
   9 & 0.2550 (0.0106) & 8.8517 (0.3200)  \\
  10 & 0.2339 (0.0178) & 9.5867 (0.6273)  \\
  11 & 0.2076 (0.0119) & 10.6565 (0.5543) \\
  12 & 0.1839 (0.0112) & 11.9082 (0.6603) \\
  13 & 0.1734 (0.0181) & 12.6228 (1.1593) \\
  14 & 0.1555 (0.0085) & 13.8911 (0.6963) \\
  15 & 0.1453 (0.0086) & 14.7963 (0.7803) \\
  16 & 0.1316 (0.0083) & 16.2468 (0.9562) \\
  17 & 0.1250 (0.0079) & 17.0421 (1.0042) \\
  18 & 0.1146 (0.0035) & 18.4593 (0.5275) \\
  19 & 0.1087 (0.0032) & 19.4153 (0.5420) \\
  20 & 0.1056 (0.0060) & 19.9855 (1.1097) \\
  21 & 0.0983 (0.0017) & 21.3548 (0.3492) \\
  22 & 0.0945 (0.0056) & 22.2248 (1.2513) \\
  23 & 0.0918 (0.0048) & 22.8346 (1.1785) \\
  24 & 0.0845 (0.0043) & 24.7050 (1.2327) \\
  25 & 0.0874 (0.0039) & 24.9125 (1.0243) \\
\end{tabular}
\caption{ POLSAR Speckle Suppression Power: ENL and Variance Log-Determinant }
\label{tab:enl_in_log_domain}
\end{table}

Other fields in variance log-det table can include: var theory, and variance analysis

\subsection{Evaluating POLSAR Speckle Filters over Homogeneous Areas}

\begin{figure}
	\subfloat[preserving consistency property: boxcar filter]{
		 \epsfxsize=3in
		 \epsfysize=3in
                 \epsffile{images/AIRSAR_Flevoland.preserve_consistency.boxcar.eps} 
		 \label{dispersion_2x2}
	} 
	\hfill	
	\subfloat[preserving consistency property: fmle filter]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/AIRSAR_Flevoland.preserve_consistency.fmle.eps} 	
		 \label{contrast_2x2}
	}      
  \caption{Checking the preservation of the consistency property for POLSAR speckle filters}  
\end{figure}

This section describes how POLSAR filters can be evaluated
  using the consistent measures of distance found in the homoskedastic log-transformed domain.
The previous section has illustrated theoretically how the variance of sample log-determinant is linked to the ENL index.
This form the basis for evaluating POLSAR speckle filters over homogeneous areas.
The procedure is simple.
To evaluate a given POLSAR speckle filter over homogeneous areas,
  the filter is applied over any known homogeneous areas and the sample variance of log-determinant is measured.
The Equivant Number of Look (ENL) is then estimated
  either by referencing the prepared graphs given by Eqn. ???
  or alternatively by setting the measured variance value into $var[\ln{|C_v|}]$ in Eqn. ???.

In order for such a procedure to generic enough, it is important that the given POLSAR speckle filter to preserve the consistency property.
That can be tested by applying the POLSAR filter into different sets of homogeneous area and investigate the plots of the dis-similarity measures presented above.
Fig. ??? presents two example plots for such a test.
Experiment is carried out for the 3x3 POLSAR boxcar filter to show that it preserves the consistency property.
The boxcar filter is applied into 2 sets of part-pol AIRSAR data over Flevoland (HH-HV and VH-VV).
Log-determinant and the contrast measure is computed for the inputs and outputs filtered POLSAR data,
  and their plots are presented in Fig. ???

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[Log-determinants histograms of boxcar 3x3 speckle filter]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/boxcar_3x3_preserves_consistency.log_determinant.eps} 	
		 \label{log_determinant}
	} 
	\hfill	
	\subfloat[Contrast histograms of boxcar 3x3 speckle filter]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/boxcar_3x3_preserves_consistency.contrast.eps} 	
		 \label{contrast}
	}   
\end{tabular}
\caption{POLSAR 3x3 boxcar filter preserves the consistency property. Consistency means: as long as the area is homogeneous, regardless of the underlysing signal $\Sigma_v$ the shapes of the histograms should be the same.}
\label{fig:boxcar_3x3_preserves_consistency}
\end{figure}

Apparently for log-determinant histograms,
  while the location of the filtered plots differs depending on the underlysing signal $\Sigma_v$,
  the shape of the plots, for both pre-filtered and post-filtered data, is invariant to that underlying signal.
Contrast, and similarly dispersion, measures of distance have its location independent of $\Sigma_v$,
  thus the contrast plots for the filtered output should overlap each other to show consistency.

The consistency property of POLSAR speckle filter is important
  not only to make the estimation of ENL become general enough
  but also to ensure that any classification / detection algorithm
    which is based on the scalar and consistent measures of distance would work on both pre-filtered and post-filtered data.

\subsection{Evaluating POLSAR Speckle Filters over Heterogeneous Areas}

Over heterogeneous area, the consistent measures of distance may also be valuable tools in helping to evaluate POLSAR speckle filters.
For a start, since the model for log-determinant is additive and homoskedastic,
  log-determinant images may be better naturally suited for gray-level digital images.
Especially for evaluation of statistical estimators,
  the estimators' error / residual plays an important role.
Ideally speaking, under the context of additive model,
  the perfect estimators' residual should be consists of only and wholly random noise.
And Gauss-Markov theorem, which works under the assumption of homoskedasticity, states that the optimal estimator should have minimal Mean Squared Error.
When the underlying signal is not known apriori,
  the second best gauge is possibly having the MSE of residual being equals or being as close as possible to the MSE of the inherent noise.

Our previous work TODO:CITE explored the hypothesis context of SAR data.
Here an experiment is carried out to illustrate the use of the consistent and homoskedastic measures of distance in evaluating the performance of boxcar 3x3 and boxcar 5x5 POLSAR filter on the AIRSAR Flevoland partial POLSAR data (HH-HV).
A square 740x740 portion of the AIRSAR dataset is extracted, and the two POLSAR speckle filters is applied.
The log-determinant images of outputs are display in Fig. ???.
For future further analysis, the residual defined as the distance between the log-determinants of the outputs and the original input is computed for both cases and the images are also displayed in the same figure.
Assumming the quantitative evaluation of SAR speckle filters can also be extended to POLSAR speckle filters,
  the Mean Squared Error (MSE) of the filters are computed.
Using the value of 4 as the ENL for the AIRSAR Flevoland dataset,
  the ``optimal'' value is computed as
%  \begin{equation}
%    %avg(\mathbb{L})^2 + var(\mathbb{L}) = mse(\mathbb{L}) =
%    \left[ \psi^0(4) + \psi^0(3) - 2\ln4 \right]^2 + \left[ \psi^1(4) + \psi^1(3) \right] = 1.0312 
%  \end{equation}
by setting $d=2,L=4$ into Eqn. (TODO:REF) making the expected MSE being $mse(\mathbb{L})=1.0132$.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[Log-determinant Image of boxcar 3x3 speckle filter]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/visual_eval_part_pol_boxcar_3.filtered.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[Log-determinant Image of boxcar 5x5 speckle filter]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/visual_eval_part_pol_boxcar_5.filtered.eps} 	
		 \label{multi_look_contrast}
	} \\
	\subfloat[Image of Log-determinant Residual for 3x3 filter (MSE=1.5594)]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/visual_eval_part_pol_boxcar_3.residual.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[Image of Log-determinant Residual for 5x5 filter (MSE=2.1420)]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/visual_eval_part_pol_boxcar_5.residual.eps} 	
		 \label{multi_look_contrast}
	} 
\end{tabular}
\caption{Visually Evaluating POLSAR Boxcar 3x3 vs. 5x5 Speckle Filters on AIRSA Flevoland part-pol data (HH-HV) with expected MSE=1.0312 at ENL=4. }
\label{fig:visual_eval_part_pol_boxcar_speckle_filters_3x3_vs_5x5}
\end{figure}
%Note if the picture does not look convincing enough
%another option here is to show 7x7 filter with more pronouced blurring MSE=2.5
Visually speaking, while the worse blurring effects of the boxcar 5x5 speckle filter is quite hard to be observed in the final image of the filtered output,
  such detection can be made relatively easier by just by visually investigating the residual image.
And when quantified number is taken out and compared with the expected level of noise to be removed,
  the excessive blurring effects of the 5x5 filter become crystally clear.
In fact, even the 3x3 boxcar filter itself might be also a bit blurry,  as suggested by its relatively high residual values.% of the 3x3 boxcar filter suggests that   
But a full explaination for that is outside the scope of this paper.

\begin{figure}
 	\subfloat[boxcar filter: log-determinant image]{
		 \epsfxsize=3in
		 \epsfysize=3in
                 \epsffile{images/AIRSAR_Flevoland.boxcar.eps} 
		 \label{dispersion_2x2}
	} 
	\hfill	
	\subfloat[boxcar filter: residual image]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/AIRSAR_Flevoland.boxcar.residual.eps} 	
		 \label{contrast_2x2}
	}  \\
 	\subfloat[fmle filter: log-determinant image]{
		 \epsfxsize=3in
		 \epsfysize=3in
                 \epsffile{images/AIRSAR_Flevoland.fmle.eps} 
		 \label{dispersion_2x2}
	} 
	\hfill	
	\subfloat[fmle filter: residual image]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/AIRSAR_Flevoland.fmle.residual.eps} 	
		 \label{contrast_2x2}
	}   
  \caption{AIRSAR image in log-determinant format and its variance as an edge-indicator}  
\end{figure}