### 2013-03-01 7/discussion.tex

% this file is called up by thesis.tex
% content in this file will be fed into the main document

\chapter{Discussion} % top level followed by section, subsection

% ----------------------- paths to graphics ------------------------

% change according to folder and file names
\ifpdf
    \graphicspath{{7/figures/PNG/}{7/figures/PDF/}{7/figures/}}
\else
    \graphicspath{{7/figures/EPS/}{7/figures/}}
\fi

% ----------------------- contents from here ------------------------

Thus log transformation is shown here to be a homomorphic transformation, allowing one to apply traditional linear, 
additive, least squared error regression signal processing (e.g. wavelets) and computational inteliggence
(e.g. artificial neural network) techniques on SAR data.

% ---------------------------------------------------------------------------
% ----------------------- end of thesis sub-document ------------------------
% ---------------------------------------------------------------------------