### 2014-03-27 answer_thesis_examiner/answer_document.tex
auto

\documentclass{article}

\usepackage{titlesec}
\newcommand{\sectionbreak}{\clearpage}

\title{Reply To Examiners}
\date{\today}
\author{Thanh-Hai Le}

\begin{document}
\maketitle

\newenvironment{replyheader}{
\begin{tabular}{l l}
\bf{Name of Student:} & Le Thanh Hai \\
\bf{Degree:} & Doctor of Philosophy \\
\bf{Thesis Title:} & Scalar \& Homoskedastic Models for SAR \& POLSAR data
\end{tabular}  
}

\section*{Reply to Examiner No. 1}

\begin{replyheader}
\end{replyheader}

\section*{Reply to Examiner No. 2}

\begin{replyheader}
\end{replyheader}

\section*{Reply to Examiner No. 3}

\begin{replyheader}
\end{replyheader}

\end{document}