### 2014-04-12 answer_thesis_examiner/answer_document.tex
auto

\documentclass{article}

\usepackage{titlesec}
\newcommand{\sectionbreak}{\clearpage} %to give each section a new page

\usepackage{amsmath} %to have equation with no number
\usepackage{chapterbib} %to have references in each chapter
%\usepackage{enumerate} %to enumerate with a,b,c instead of 1,2,3
\renewcommand{\theenumi}{\alph{enumi}}%to enumerate with a,b,c instead of 1,2,3

\usepackage{enumitem}
%\setlist{nolistsep} % no space among items as well as around whole list
\setlist{noitemsep} %leave space around whole list

\title{Reply To Examiners}
\date{\today}
\author{Thanh-Hai Le}

\begin{document}
\maketitle

\newenvironment{replyheader}{
	\vspace*{10mm}
\setcounter{page}{1}

\begin{tabular}{l l}
\bf{Name of Student:} & Le Thanh Hai \\
\bf{Degree:} & Doctor of Philosophy \\
\bf{Thesis Title:} & Scalar \& Homoskedastic Models \\ 
 & for SAR \& POLSAR data
\end{tabular}  
	\vspace*{10mm}
}

\include{chap1}

\include{chap2}

\include{chap3}

\end{document}