### 2014-04-24 answer_thesis_examiner/old.chap2.tex
auto commit

\section*{Reply to Examiner No. 2}

\begin{replyheader}
\end{replyheader}

Examiner No. 2 suggests to further emphasize the significance of the research motivation in \textbf{Section 1.1 and Section 1.2}.
In a similar fashion, the conclusion chapter should also include the research objectives.
These have been updated in the revised thesis.
Even though I would wish to note that the research objectives have been spelled out in the Introduction chapter.
The examiner also suggested to change the word ``theory'' to ``model''.
That has also been applied through out the thesis.

\textbf{The second main comment} is to further highlight the novelty claim, which can be splitted into two parts.

\textit{In the first part}, the examiner appears to discount the novelty claim by suggesting that:
 ``the proposed model is derived based on the existing statistical models for SAR and extended to POLSAR''.
The main approach, in actuality, differs in a subtle but important way:
The proposed model is derived based on generic mathematical results for multi-dimensional random-walk \cite{Goodman_JOptSocAm_76, Goodman_Springer_1975}.
These results are well known to be applicable to POLSAR.
Thus, the derived models by nature are multi-dimensional model.
However, when collapsed into single-dimension, they match perfectly with existing SAR models.
As such, the proposed models can be considered as an extension of existing statistical models for SAR.

\textit{In the second part}, the Examiner expresses his doubt in ``the use of MSE (mean squared error) as the single unified evaluation criteria''.
His main concern: ``the proposed log-transform model will introduce an inevitable bias error''.
It is no doubt tricky to evaluate evaluation criteria.
And I can empathize with his feelings.
Still, I wish to point out that MSE evaluation inherently includes bias evaluation. 
In fact it has two components: 1. bias evaluation and 2. variance evaluation, which for SAR speckle filter evaluation translates into 1. radiometric preservation and 2. noise suppression respectively.

%( \textbf{ Hai note: }  The Examiner mentions further clarifications to support the claim of novelty in Part II of his report.
%Unfortunately I do not have it at hand.
%It would be great if I can have a look at it!)

The Examiner also suggestion the inclusion of several particular references into \textbf{the literature review}.
I really appreciate the examiner effort in pointing out specific references \cite{Argenti_GRSM_2013, Lee_RSReviews_1994, Cetin_ProcSPIE_2000, White_ProcSPIE_1994, Sattar_TIP_1997, Wang_TIP_2004, Nielsen_2012_ICASSP}.
They, and some other articles, have been included in the thesis.

%The examiner also suggestion inclusion of several references into \textbf{the literature review} topics of ``relevant SAR/POLSAR speckle filters'' and  ``evaluation of SAR speckle filters''.
%I really appreciate the examiner effort in pointing out specific references.
%They, and some other articles, have been included in the thesis.
%I would like however to seek his empathy towards a couple of points.
%First, a number of the references cited has already been included into the original thesis (e.g \cite{Touzi_2002_TGRS} or \cite{Xie_2002_TGRS}.). 
%Some of them (for example \cite{Argenti_GRSM_2013}) are actually published (Sept 2013) after the submission of the original thesis (Aug 2013).
%Second, please also note that the topics of SAR/POLSAR speckle filters and to some extend, their evaluation, may be considered as part of the thesis's contribution.
%While they do results in published scientific articles, 
%  they are not, however, to be considered as the main contributions of the thesis.
%Rather they serves as application, or demonstration, of the benefits of the thesis's main proposal.
%As such, while I understand the need for comprehensive review, due to various constraint, I seek his understanding that they are probably not to be explored with the same depth and coverage as the main topic of the thesis.

\textbf{To conclude}, I appreciate the fact that the Examiner is able to a summarize the long thesis into a single sentence.
Also, portions of this thesis has been extracted out as scientific articles and submitted for peer review.
Various errata exposed by Examiner No. 2 and have also been updated in the thesis.
Last but not least, I wish to express my appreciation towards Examiner No.2 for his careful and detailed review.

%\bibliographystyle{plainnat}
\bibliographystyle{apalike}
\bibliography{answer_document}